---
title: "Comparing Two Proportions"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(openintro)
library(tidyverse)
library(infer)

fish_oil <- fish_oil_18[['major_cardio_event']]

fish_oil_full_data <- tibble(treatment = c(rep("fish_oil", 386), rep("placebo", 419), rep("fish_oil", 12547), rep("placebo", 12519)),
                             outcome = c(rep("major_cardio_event", 386+419), rep("no_event", 12547+12519)))

p_fo <- 386/(386+12547)
p_p <- 419/(419+12519)


p_diff <- p_fo - p_p

p_pool <- (39+26)/90

SE <- sqrt(p_fo*(1-p_fo)/(386+12547) + p_p*(1-p_p)/(419+12519) )

p_T <- 26/(26+14)
p_C <- 39/(39+11)
  
nT <- (26+14)
nC <- (29+11)
  
z <- (p_T- p_C) / sqrt(p_pool*(1-p_pool)*(1/nT+1/nC))

z_star <- qnorm(.975, mean=0, sd=1)

knitr::opts_chunk$set(echo = FALSE)
```


## Fish Oil on Heart Attacks

A 5-year experiment was conducted to evaluate the effectiveness of fish oils on reducing heart attacks, where each subject was randomized into one of two treatment groups. We’ll consider the data stored in the contingency table `fish_oil`. Take a look around the data set below. 

```{r sandbox, exercise=TRUE}
fish_oil
```

Suppose you are interested in estimating the difference in the rate of major cardiac events between those in the `fish_oil` and `placebo` groups.

```{r ht-v-ci-1}
question_radio(
"Which of the following is the most appropriate course of action?",
answer("Construct a 95% confidence interval", correct=TRUE),
answer("Conduct a hypothesis test"),
allow_retry=TRUE
)
```

Your task is construct a 95% confidence interval for the difference in the rate of major cardiac events between those in the `fish_oil` and `placebo` groups. First, verify that our data meets the success failure condition:

```{r success-failure-1, exercise=TRUE}

```

```{r success-failure-1-hint-1, exercise=TRUE}
# by looking at the contingency table stored in fish_oil we can see that both
# the fish_oil and placebo groups have at least 10 successes and 10 failures
```


If you feel up to it, try and construct the confidence interval console below. Note that since our data is in the form of a contingency table, we won't be able to use the `infer` package. If you don't feel comfortable doing this the remaining consoles will walk you through the steps.

```{r ci-advanced, exercise=TRUE}

```

Remember that confidence intervals take the form *point estimate* $\pm z^*\times SE$. Let first compute $\hat{p}_1-\hat{p}_2$, our point estimate:

```{r point_est, exercise=TRUE}

```

```{r point_est-hint-1}
p_fo <- _____
p_p <- _____

p_diff <- ______ - ______

p_fo
p_p
p_diff
```

```{r point_est-hint-2}
p_fo <- _____
p_p <- _____

p_diff <- p_fo - p_p

p_fo
p_p
p_diff
```

```{r point_est-hint-3}
p_fo <- 386/(386+12547)
p_p <- 419/(419+12519)


p_diff <- p_fo - p_p

p_fo
p_p
p_diff
```

Now that we have our point estimate, compute the standard error (technically the estimate of the standard error):

```{r standard-err, exercise=TRUE}

```

```{r standard-err-hint-1}
SE <- sqrt(_____*(1-_____)/_____ + _____*(1-_____)/_____ )
SE
```

```{r standard-err-hint-2}
SE <- sqrt(p_fo*(1-p_fo)/(386+12547) + _____*(1-_____)/_____ )
SE
```

```{r standard-err-hint-3}
SE <- sqrt(p_fo*(1-p_fo)/(386+12547) + p_p*(1-p_p)/(419+12519) )
SE
```

Compute $z^*$ for a 95% confidence interval:

```{r z-star, exercise=TRUE}

```

```{r z-star-hint-1}
z_star <- qnorm(_____, _____, _____)
z_star
```

```{r z-star-hint-2}
z_star <- qnorm(_____, mean=_____, sd=_____)
z_star
```

```{r z-star-hint-3}
z_star <- qnorm(.975, mean=0, sd=1)
z_star
```

Now that we have all the ingredients that we need, compute the confidence interval:

```{r con-int, exercise=TRUE}

```

```{r con-int-hint-1}
_____ - _____ * _____
_____ + _____ * _____
```

```{r con-int-hint-2}
p_diff - z_star * SE
p_diff + z_star * SE
```


So our 95% confidence interval is (-.68%, .17%).

Now let's repeat this analysis using the `infer` package. I have created a tibble called `fish_oil_full_data` which can be used. Follow the following steps to construct a 95\% confidence interval:

1. Define the sampling distribution. 
2. Compute the point estimate.
3. Compute the confidence interval.
4. (Optional) Visualize sampling distribution and confidence interval.

```{r infer-ci, exercise=TRUE}

```

```{r infer-ci-hint-1}
samp_dist <- fish_oil_full_data %>% 
  # specify the variable you're interested in
  specify(_____, _____, _____) %>% 
  # specify sampling distribution
  assume(______)

point_est <- fish_oil_full_data %>% 
  # specify the variable you're interested in
  specify(______, _____, _____) %>% 
  # specify the statistic you want to compute
  calculate(_____, _____)

point_est

ci <- samp_dist %>% 
  get_ci(_____, _____)

ci

samp_dist %>% 
  visualize() +
    shade_confidence_interval(_____) 
```

```{r infer-ci-hint-2}
samp_dist <- fish_oil_full_data %>% 
  # specify the variable you're interested in
  specify(response = _____, explanatory = _____, success = "______") %>% 
  # specify sampling distribution
  assume(distribution = "_____")

point_est <- fish_oil_full_data %>% 
  # specify the variable you're interested in
  specify(response = _____, explanatory = _____, success = "_____") %>% 
  # specify the statistic you want to compute
  calculate("______", order = c("_____", "_____"))

point_est

ci <- samp_dist %>% 
  get_ci(level = _____, point_estimate = _____)

ci

samp_dist %>% 
  visualize() +
    shade_confidence_interval(_____) 
```

```{r infer-ci-hint-3}
samp_dist <- fish_oil_full_data %>% 
  # specify the variables you're interested in
  specify(response = outcome, explanatory = treatment, success = "major_cardio_event") %>% 
  # specify sampling distribution
  assume(distribution = "z")

point_est <- fish_oil_full_data %>% 
  # specify the variables you're interested in
  specify(response = outcome, explanatory = treatment, success = "major_cardio_event") %>%
  # specify the statistic you want to compute
  calculate("diff in props", order = c("fish_oil", "placebo"))

point_est

ci <- samp_dist %>% 
  get_ci(level = 0.95, point_estimate = point_est)

ci

samp_dist %>% 
  visualize() +
    shade_confidence_interval(ci)
```


## CPR and Blood Thinners

An experiment consisted of two treatments on patients who underwent CPR for a heart attack and were subsequently admitted to a hospital. Each patient was randomly assigned to either receive a blood thinner (treatment group) or not receive a blood thinner (control group). The outcome variable of interest was whether the patient survived for at least 24 hours. The data is stored in `cpr`:

```{r sandy, exercise=TRUE}

```


```{r ht-v-ci-2}
question_radio(
"If you'd like to determine whether giving patients a blood thinner impacts mortality, what is the best course of action?",
answer("Construct a 95% confidence interval"),
answer("Conduct a hypothesis test", correct=TRUE),
allow_retry=TRUE
)
```


Your task is to test whether giving patients a blood thinner impacts mortality (either increasing or decreasing). If you feel up to it, try and do it all in the console below. This includes verifying our data satisfies the success-failure criterion and selecting your own significance level. I'd like you to conduct the hypothesis test both with and without the `infer` package. If you don't feel comfortable doing this the remaining consoles will walk you through the steps.

```{r advanced, exercise=TRUE}

```


Call the mortality rate of the population that receives the treatment $p_T$ and the mortality rate of those who do not $p_C$. On a sheet of paper or in the console, write down what you think your null and alternative hypotheses are:

```{r hypotheses, exercise=TRUE}
# H0:
# HA:
```


```{r hypotheses-solution}
# H0: p_T - p_C = 0
# HA: p_T - p_C ≄ 0
```

Create a contingency table:

```{r cont_tab, exercise=TRUE}

```

```{r cont_tab-hint-1}
______ %>% 
  _____ %>% 
  ______
```

```{r cont_tab-hint-2}
cpr %>% 
  count(_____, ______) %>% 
  pivot_wider(names_from=______, values_from = _____)
```

```{r cont_tab-hint-3}
cpr %>% 
  count(group, outcome) %>% 
  pivot_wider(names_from=group, values_from = n)
```

Compute $\hat{p}_{\textit{pool}}$

```{r pool, exercise=TRUE}

```

```{r pool-hint-1}
p_pool <- ____/______
```

```{r pool-hint-2}
p_pool <- (39+26)/90
p_pool
```


Does this test/data meet the success-failure criterion?

```{r s-f, exercise=TRUE}

```

```{r s-f-hint-1}
p_pool * ____
(1-p_pool) * ____
p_pool * ____
(1-p_pool) * ____
```

```{r s-f-hint-2}
p_pool * (39+11)
(1-p_pool) * (39+11)
p_pool * (26+14)
(1-p_pool) * (26+14)

# Since all these values are above 10 we satisfy the success-failure criterion
```

Compute the test statistic (z-score). It shouldn't matter which order you do your subtraction but let's do $p_T-p_C$:

```{r test-statistic, exercise=TRUE}

```

```{r test-statistic-hint-1}
p_T <- _____
p_C <- _____
  
nT <- _____
nC <- _____
  
z <- _____
z
```

```{r test-statistic-hint-2}
p_T <- 26/(26+14)
p_C <- 39/(39+11)
  
nT <- (26+14)
nC <- (29+11)
  
z <- _____
z
```

```{r test-statistic-hint-3}
p_T <- 26/(26+14)
p_C <- 39/(39+11)
  
nT <- (26+14)
nC <- (29+11)
  
z <- (p_T- p_C) / sqrt(p_pool*(1-p_pool)*(1/nT+1/nC))
z
```

Use the `normTail` function to display the region of the standard normal distribution corresponding to your p-value:

```{r z, exercise=TRUE}

```

```{r z-hint-1}
normTail(____, ____, _____, ____)
```

```{r z-hint-2}
normTail(m=_____, s=____, L=_____, U=_____)
```

```{r z-hint-3}

normTail(m=0, s=1,  L=_____, U=_____)
```

```{r z-hint-4}
normTail(m=0, s=1, L=z, U=-z)
```

Compute your p-value:

```{r p-val, exercise=TRUE}

```

```{r p-val-hint-1}
2*pnorm(____, ____, ____)
```

```{r p-val-hint-2}
2*pnorm(____, mean = ____, sd = ____)
```

```{r p-val-hint-3}
2*pnorm(z, mean = 0, sd = 1)
```

```{r q-1}
question_radio(
"Would is the result of this hypothesis test with a significance level of $\\alpha = 0.05$",
answer("Fail to Reject the Null Hypothesis", correct=TRUE),
answer("Reject the null hypothesis"),
allow_retry=TRUE
)
```

```{r q-4}
question_radio(
"What is an appropraite interpretation of this result?",
answer("There is no impact of blood thinners on the efficacy of CPR."),
answer("Blood thinners are an effective way of improving CPR"),
answer("Blood thinners decrease the efficacy of CPR"),
answer("We don't have enough evidence to conclude that blood thinners have an impact on CPR.", correct = TRUE),
allow_retry=TRUE
)
```

Now we'll repeat the process using the `infer` package. In the console below:

1. Define the null distribution.
2. (Optional) Visualize the null distribution.
3. Compute the test statistic.
4. (Optional) Visualize the p-value.
5. Compute the p_value.

```{r infer-1, exercise=TRUE}

```

```{r infer-1-hint-1}
null_dist <- cpr %>% 
  specify_____, _____, _____) %>% 
  hypothesize(_____) %>% 
  assume(_____)

null_dist %>% 
  visualize()

test_stat <- cpr %>% 
  specify_____, _____, _____) %>% 
  hypothesize(______) %>% 
  calculate(_____, _____)

null_dist %>% 
  visualize() +
    shade_p_value(______, ______)

null_dist %>% 
  get_p_value(_____, _____)
```

```{r infer-1-hint-2}
null_dist <- cpr %>% 
  specify(response = _____, explanatory = _____, success = "_____") %>% 
  hypothesize(null = "_____") %>% 
  assume(distribution = "_____")

null_dist %>% 
  visualize()

test_stat <- cpr %>% 
  specify(response = ______, explanatory = _____, success = "_____") %>% 
  hypothesize(null = "_____") %>% 
  calculate(stat = "_____", order = c("_____", "_____"))

null_dist %>% 
  visualize() +
    shade_p_value(obs_stat = _____, direction = "______")

null_dist %>% 
  get_p_value(obs_stat = ______, direction = "______")
```

```{r infer-1-hint-3}
null_dist <- cpr %>% 
  specify(response = outcome, explanatory = group, success = "died") %>% 
  hypothesize(null = "independence") %>% 
  assume(distribution = "z")

null_dist %>% 
  visualize()

test_stat <- cpr %>% 
  specify(response = outcome, explanatory = group, success = "died") %>% 
  hypothesize(null = "independence") %>% 
  calculate(stat = "z", order = c("treatment", "control"))

null_dist %>% 
  visualize() +
    shade_p_value(obs_stat = test_stat, direction = "two-sided")

null_dist %>% 
  get_p_value(obs_stat = test_stat, direction = "two-sided")
```

