---
title: "Introduction To Data"
output: 
  learnr::tutorial:
    progressive: false
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(openintro)
library(usdata)
knitr::opts_chunk$set(echo = FALSE)
```

## Data Basics

### Observations, variables, and data matrices

In the console below use `?` to find the help page for the function `head`.

```{r head-help, exercise=TRUE}

```

```{r head-help-hint}
?head
```

Use the `head` function to display the first 6 rows of the `county` dataset:

```{r display-county, exercise=TRUE}

```

```{r display-county-hint-1}
head(____)
```

```{r display-county-hint-2}
head(county)
```

Now let's review some vocabulary:

* **case** or **observational unit**: each row, a single unit of data
* **variable**: each column, a property of each unit of data
* **data frame**: whole table
* **tidy data**: data frame where each row is a unique case (observational unit), each column is a variable, and each cell is a single value

When working with data, it's extremely important that we understand important information about our data, such as what each variable means and what it's unit of measurement is. In the code box below type `?` before `county` to get the **metadata** for this  dataset:

```{r county-meta, exercise=TRUE}

```

```{r county-meta-hint}
?county
```

Use the code box below to figure out how many *cases* and how many *variables* are in the `county` dataset:

```{r county-sandbox, exercise=TRUE}

```

<div id="county-sandbox-hint">
**Hint:** You may want to use the `glimpse` function.
</div>

```{r cases-question, echo=FALSE}
question("How many *cases* are in the `county` dataset?",
  answer(2000),
  answer(3142, correct = TRUE),
  answer(15),
  answer(43671)
)
```

```{r variables-question, echo=FALSE}
question("How many *variables* are in the `county` dataset?",
  answer(2000),
  answer(3142),
  answer(15, correct = TRUE),
  answer(43671)
)
```


### Types of variables

Recall the different types of variables:

* **numerical**: takes range of numerical values, and it is sensible to add, subtract, or take averages with those values
  + **discrete**: only takes numerical values in jumps
  + **continuous**: takes all values in an interval
* **categorical**: values are categories called **levels**
  + **ordinal**: the **levels** have a natural ordering
  + **nominal**: the **levels** have no ordering
  
Use the following sandbox to answer the following questions about the dataset `county`:
```{r county-sandbox2, exercise=TRUE}

```


```{r variables-question1, echo=FALSE}
question("What kind of variable is `name`?",
  answer("Numerical, Continuous"),
  answer("Numerical, Discrete"),
  answer("Categorical, Ordinal"),
  answer("Categorical, Nominal", correct = TRUE),
  allow_retry = TRUE
)
```

```{r variables-question2, echo=FALSE}
question("What kind of variable is `pop_change`?",
  answer("Numerical, Continuous", correct = TRUE),
  answer("Numerical, Discrete"),
  answer("Categorical, Ordinal"),
  answer("Categorical, Nominal"),
  allow_retry = TRUE
)
```

```{r variables-question3, echo=FALSE}
question("What kind of variable is `smoking_ban`?",
  answer("Numerical, Continuous"),
  answer("Numerical, Discrete"),
  answer("Categorical, Ordinal", correct = TRUE),
  answer("Categorical, Nominal"),
  allow_retry = TRUE
)
```

```{r variables-question4, echo=FALSE}
question("What kind of variable is `multi_unit`?",
  answer("Numerical, Continuous", correct = TRUE),
  answer("Numerical, Discrete"),
  answer("Categorical, Ordinal"),
  answer("Categorical, Nominal"),
  allow_retry = TRUE
)
```

### Relationships between variables
Recall the following definitions:

* **positive association**: when one variable increases, the other variable also increases
* **negative association**: when one variable increases, the other variable decreases
* **independent**: when two variables are not associated.


In the console below plot any two variables which are positively associated in the `county` data set.

```{r county-positive-association, exercise=TRUE}

```

```{r county-positive-association-hint}
ggplot(____, aes(x=____, y=____)) +
  geom_point()
```

In the console below plot any two variables which are negatively associated in the `county` data set.

```{r county-negative-association, exercise=TRUE}

```

### Explanatory and Response Variables

When we suspect one variable might causally affect another, we label the first variable the **explanatory variable** and the second the **response variable**. We also use the terms explanatory and response to describe variables where the response might be predicted using the explanatory even if there is no causal relationship.

In console below, write down a research question about the relationship of two numerical variables in the `county` dataset. Write this as a comment. Remember you can use the `#` symbol to write comments in `R`. What do you expect the relationship between these variables to be? What are some confounding variables that may impact the relationship between these two variables?

```{r research-question, exercise=TRUE}

```

What is the explanatory and response variable in your research question?

```{r research-exp-resp, exercise=TRUE}

```

Plot the explanatory vs. the response variable below.

```{r research-scatter, exercise=TRUE}

```

Describe the relationshipo between your explanatory and response variables in
the context of your research question. Be prepared to discuss in class.

