---
title: "19 Inference For A Single Mean"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(openintro)
library(tidyverse)
library(infer)

set.seed(1989)

mankato <- read_csv("https://raw.githubusercontent.com/EricFriedlander/MATH221SP22/main/Datasets/mankato_flood.csv")

benzene <- tibble(level = rnorm(20, 0.0055, sd = 0.002))


knitr::opts_chunk$set(echo = FALSE)
```


## Mankato Flooding

In this activity we'll continue the example we worked on in the lecture. This time we'll create a 99% confidence interval for the elevation of the flood waters. This variable is called `Stage` and is stored in the `mankato` data frame:

```{r sandbox-1, exercise=TRUE}

```


If you feel up to it, please create the 99% confidence interval in the console below. This includes doing a little exploratory data analysis a verifying that the data is suitable for the CLT. If you don't feel comfortable doing this the remaining consoles will walk you through the steps.

```{r self-1, exercise=TRUE}

```

In the console below generate a histogram and box plot of the `Stage` variable. For the histogram, use 8 bins:

```{r stage-viz, exercise=TRUE}

```

```{r stage-viz-hint-1}
ggplot(_____) +
  _____(_____)

ggplot(_____) +
  _____(_____)
```

```{r stage-viz-hint-2}
ggplot(_____, _____) +
  geom_histogram(_____)

ggplot(_____, _____) +
  geom_boxplot()
```

```{r stage-viz-hint-3}
ggplot(_____, aes(x=_____)) +
  geom_histogram(bins=_____)

ggplot(_____, aes(x=_____)) +
  geom_boxplot()
```

```{r stage-viz-hint-4}
ggplot(mankato, aes(x=Stage)) +
  geom_histogram(bins=8)

ggplot(mankato, aes(x=Stage)) +
  geom_boxplot()
```

```{r q-1}
question_radio("(Practice Test Question): Does this data seem to meet the conditions for the central limit theorem? On a separate sheet of paper explain your answer.",
               answer("Yes", correct=TRUE),
               answer("No"),
               allow_retry = TRUE)
```

The data has little to no skew and there are not outliers. In addition, the sample size is 85 which is quite large so we should feel comfortable using the central limit theorem.

Recall the the formula for a confidence interval is $\bar{x} \pm t^* SE$ where $SE = s/\sqrt{n}$. In the console below, compute $\bar{x}$ and $SE$:

```{r xbar-se-1, exercise=TRUE}

```

```{r xbar-se-1-hint-1}
mankato %>% 
  _____
```

```{r xbar-se-1-hint-2}
mankato %>% 
  summarize(_____)
```

```{r xbar-se-1-hint-3}
mankato %>% 
  summarize(xbar = _____,
            se = _____)
```

```{r xbar-se-1-hint-4}
mankato %>% 
  summarize(xbar = mean(_____),
            se = sd(_____)/sqrt(_____))
```

```{r xbar-se-1-hint-5}
mankato %>% 
  summarize(xbar = mean(Stage),
            se = sd(Stage)/sqrt(n()))
```


Now let's find $t^\star$ for a 99% confidence interval.

```{r q-2}
question_numeric("How many degrees of free would you use in this t-distribution?",
                 answer(84, correct=TRUE),
                 allow_retry = TRUE)
```

In the console below, find $t^\star$:

```{r t-1, exercise=TRUE}

```

```{r t-1-hint-1}
qt(_____, df=______)
```

```{r t-1-hint-2}
qt(.995, df=85)
```

In the console below, compute the final confidence interval:

```{r ci-1, exercise=TRUE}

```


```{r ci-1-hint-1}
_____ - _____ * _____

_____ + _____ * _____
```

```{r ci-1-hint-2}
763.6974	 - 2.634914 * 0.600952

763.6974	 + 2.634914 * 0.600952
```

So our point estimate for the average Stage of the maximum flood is 763.6974 feet with 99% confidence interval of (762.1139, 765.2809) feet.


## Benzene in drinking water

The Environmental Protection Agency specifies that amount of a variety of chemicals that are safe to be in our drinking water. You can see a list [here](https://www.epa.gov/ground-water-and-drinking-water/national-primary-drinking-water-regulations#one). Today we're going to be focusing on Benzene. The maximum allowable about of Benzene in our drinking water is 0.005 mg/L. Suppose you work at a water treatment plant and, instead of relying on a single measurement, you take 20 different measurements of the drinking water. The data from the most recent set of tests is stored in the data frame `benzene`. Take a look:

```{r sandbox-2, exercise=TRUE}

```

Your task is the perform a hypothesis test to see if that the concentration of benzene in the water is above 0.005 mg/L. If you feel comfortable doing this on you own please do so in the console below. Don't forget to check whether the data meets the normality assumption:

```{r self-2, exercise=TRUE}

```

In the console below or on a separate sheet of paper write down the null and alternative hypotheses for this test:

```{r hyp, exercise=TRUE}
#
```

```{r hyp-solution}
# H0: mu <= 0.005 (<= mean less than or equal to)

# HA: me > 0.005
```


In the console below, create a histogram (use 6 bins) and box plot of the data:

```{r ben-viz, exercise=TRUE}

```

```{r ben-viz-hint-1}
ggplot(_____) +
  _____(_____)

ggplot(_____) +
  _____(_____)
```

```{r ben-viz-hint-2}
ggplot(_____, _____) +
  geom_histogram(_____)

ggplot(_____, _____) +
  geom_boxplot()
```

```{r ben-viz-hint-3}
ggplot(_____, aes(x=_____)) +
  geom_histogram(bins=_____)

ggplot(_____, aes(x=_____)) +
  geom_boxplot()
```

```{r ben-viz-hint-4}
ggplot(benzene, aes(x=level)) +
  geom_histogram(bins=6)

ggplot(benzene, aes(x=level)) +
  geom_boxplot()
```

Recall that the formula for the t-statistics i $\frac{\bar{x}-\mu}{SE}$ wher $\mu = 0.005$ in this case. In the console below compute the t-statistic of the data. Note:

```{r t-stat, exercise=TRUE}

```

```{r t-stat-hint-1}
benzene %>% 
  _____
```

```{r t-stat-hint-2}
benzene %>% 
  summarize(_____)
```

```{r t-stat-hint-3}
benzene %>% 
  summarize(xbar = _____,
            se = _____,
            t_stat = _____)
```

```{r t-stat-hint-4}
benzene %>% 
  summarize(xbar = mean(_____),
            se = sd(_____)/sqrt(_____),
            t_stat = (_____ - _____)/_____)
```

```{r t-stat-hint-5}
benzene %>% 
  summarize(xbar = mean(level),
            se = sd(level)/sqrt(85),
            t_stat = (xbar - 0.005)/se)
```

```{r df-q}
question_numeric("When computing our p-value, how many degrees of freedom should we use?",
                 answer(19, correct=TRUE),
                 allow_retry = TRUE)
```

In the console below, compute the p-value for this test. On a separate sheet of paper sketch the t-distribution and shade the region corresponding to the p-value:

```{r p-val, exercise=TRUE}

```

```{r p-val-hint-1}
1-pt(_____, df=_____)
```

```{r p-val-hint-2}
1-pt(2.091311, df=19)
```

```{r outcome-1}
question_radio("What would be the outcome of the test with a significance level of $\alpha=0.05?",
               answer("Fail to reject"),
               answer("Reject the null", correct=TRUE),
               allow_retry = TRUE)
```

```{r outcome-2}
question_radio("What would be the outcome of the test with a significance level of $\alpha=0.01?",
               answer("Fail to reject", correct=TRUE),
               answer("Reject the null"),
               allow_retry = TRUE)
```

In the context of this problem, what do you think would be an appropriate $\alpha$ to select? Why?

## Statistics vs. parameters: one mean.

Each of the following scenarios were set up to assess an average value.  For each one, identify, in words: the statistic and the parameter.
    
a.  Georgianna samples 20 children from a particular city and measures how many years they have each been playing piano.
    
b.  Traffic police officers (who are regularly exposed to lead from automobile exhaust) had their lead levels measured in their blood.

```{r q-s, exercise=TRUE}
# a)

# b)
```

```{r q-s-solution}
# a) Statistic = average number of years the 20 students have played the piano; Parameter = average number of years of piano for all kids in the town.

# b) Statistic = average blood lead level of the traffic police officers in the study; Parameter = average blood lead level of all traffic police officers (possibly restricted to that city / area).
```


## Heights of adults, standard error.

Heights of 507 physically active individuals have a mean of 171 centimeters and a standard deviation of 9.4 centimeters.
Provide an estimate for the standard error of the mean for samples of following sizes.

a. n = 10
     
b. n = 50
     
c. n = 100
     
d. n = 1000
     
e. The standard error of the mean is a number which describes what?

```{r q-h, exercise=TRUE}
# a)

# b)

# c)

# d)

# e)
```

```{r q-h-solution}
# a) SE(Xbar) =  2.97cm.

# b) SE(Xbar) = 1.33cm.

# c) SE(Xbar) = 0.94cm.

# d) SE(Xbar) = 0.3cm.

# e) The SE(Xbar) describes how Xbar would vary from sample to sample if many different datasets were collected and the sample mean was calculated on each one.
```

## Identify the critical $t$.

A random sample is selected from an approximately normal population with unknown standard deviation.
Find the degrees of freedom and the critical $t$-value (t$^\star$) for the given sample size and confidence level.

a.  $n = 6$, CL = 90%

b.  $n = 21$, CL = 98%

c.  $n = 29$, CL = 95%

d.  $n = 12$, CL = 99%

```{r q-i, exercise=TRUE}
# a)

# b)

# c)

# d)
```

```{r q-i-solution}
# a) n=6, CL = 90%, df=6−1=5, t⋆5=2.02

# b) n=21, CL = 98%, df=21−1=20, t⋆20=2.53

# c) n=29, CL = 95%, df=29−1=28, t⋆28=2.05

# d) n=12, CL = 99%, df=12−1=11, t⋆11=3.11
```


## Find the p-value, I.
A random sample is selected from an approximately normal population with an unknown standard deviation. 
Find the p-value for the given sample size and test statistic. 
Also determine if the null hypothesis would be rejected at $\alpha = 0.05$.

a.  $n = 11$, $T = 1.91$

b.  $n = 17$, $T = -3.45$

c.  $n = 7$, $T = 0.83$

d.  $n = 28$, $T = 2.13$

```{r q-f, exercise=TRUE}
# a)

# b)

# c)

# d)
```

```{r q-f-solution}
# a) n=11 , T=1.91, df=11−1=10, p−value=0.085, Do not reject H0.
# b) n=17, T=−3.45, df=17−1=16, p−value=0.003, Reject H0.
# c) n=7, T=0.83, df=7−1=6, p−value=0.438, Do not reject H0.
# d) n=28, T=2.13, df=28−1=27, p−value=0.042, Reject H0.
```


## Interpreting confidence intervals for population mean.

For each of the following statements, indicate if they are a true or false interpretation of the confidence interval.
If false, provide a reason or correction to the misinterpretation.
You collect a large sample and calculate a 95% confidence interval for the average number of cans of sodas consumed annually per adult in the US to be (440 cans, 520 cans), i.e., on average, adults in the US consume just under two cans of soda per day.

a. 95% of adults in the US consume between 440 and 520 cans of soda per year.

b. There is a 95% probability that the true population average per adult yearly soda consumption is between 440 and 520 cans.

c. The true population average per adult yearly soda consumption is between 440 and 520 cans, with 95% confidence.

d. The average soda consumption of the people who were sampled is between 440 and 520 cans of soda per year, with 95% confidence.

```{r q-in, exercise=TRUE}
# a)

# b)

# c)

# d)
```

```{r q-in-solution}
# a) False. The interval is for the parameter (a number which describes the population), not for individual observational units.

# b) False. Although unknown, the parameter is either in the interval or it is not (so either with probability zero or probability one). There is no randomness associated with the parameter or the endpoints.

# c) True.

# d) False. The sample mean is always inside the interval (it is the center!). We are 100% confident that the sample mean is in the interval.
```


## Working backwards, II. 

A 90% confidence interval for a population mean is (65, 77). The population distribution is approximately normal and the population standard deviation is unknown. This confidence interval is based on a simple random sample of 25 observations. Calculate the sample mean, the margin of error, and the sample standard deviation. Assume that all conditions necessary for inference are satisfied. Use the 
t-distribution in any calculations.

```{r q-w, exercise=TRUE}
# 
```

```{r q-w-solution}
# The sample mean is the mid-point of the confidence interval, i.e., the average of the upper and lower bounds:
# xbar =(65+772)/2=71
# The margin of error is 77−71=6. Since n=25, df=24, and the critical t-score is t=1.71. Then, 6=1.71*s/sqrt(25)≈17.54.
```

    
## Play the piano.

Georgianna claims that in a small city renowned for its music school, the average child takes less than 5 years of piano lessons. 
We have a random sample of 20 children from the city, with a mean of 4.6 years of piano lessons and a standard deviation of 2.2 years.

a. Evaluate Georgianna's claim (or that the opposite might be true) using a hypothesis test.

b. Construct a 95% confidence interval for the number of years students in this city take piano lessons, and interpret it in context of the data.

c. Do your results from the hypothesis test and the confidence interval agree? Explain your reasoning.

```{r q-pl, exercise=TRUE}
# a)

# b)

# c)

# d)
```

```{r q-pl-solution}
# a) We will conduct a 1-sample t-test. H0: μ=5. HA: μ≠5. We’ll use α=0.05. This is a random sample, so the observations are independent. To proceed, we assume the distribution of years of piano lessons is approximately normal. SE=2.2/sqrt(20)=0.4919. The test statistic is T=(4.6−5)/SE=−0.81. df=20−1=19. The one-tail area is about 0.21, so the p-value is about 0.42, which is bigger than α=0.05 and we do not reject H0. That is, we do not have sufficiently strong evidence to reject the notion that the average is 5 years.

# b) Using SE=0.4919 and t⋆df=19=2.093, the confidence interval is (3.57, 5.63). We are 95% confident that the average number of years a child takes piano lessons in this city is 3.57 to 5.63 years.

# c) They agree, since we did not reject the null hypothesis and the null value of 5 was in the t-interval.
```


## Submit

```{css echo=FALSE}
@media print {
  .topicsContainer,
  .topicActions,
  .exerciseActions .skip {
    display: none;
  }
  .topics .tutorialTitle,
  .topics .section.level2,
  .topics .section.level3:not(.hide) {
    display: block;
  }
  .topics {
    width: 100%;
  }
  .tutorial-exercise, .tutorial-question {
    page-break-inside: avoid;
  }
  .section.level3.done h3 {
    padding-left: 0;
    background-image: none;
  }
  .topics .showSkip .exerciseActions::before {
    content: "Topic not yet completed...";
    font-style: italic;
  }
}
```

Make sure you have the tutorial open in a browser (preferably Google Chrome). Please press `print page` button below to print the tutorial. Print it to a `pdf` file and upload it into Moodle. Make sure all code boxed are visible. If you can't do that please submit your answers to [the Google Form](https://forms.gle/orY1JnvpxZGiwDZP7).

```{js print2pdf1, context="server"}
    // the following 2 chunks print the completed sections of the tutorial to PDF
    // uses "css/print2pdf.css"
    $(document).on('shiny:inputchanged', function(event) {
      if (event.name === 'print2pdf') {
        window.print();
      }
    });
```
    
```{r print2pdf2}
    # button can be placed anywhere in the tutorial
    actionButton("print2pdf", "Print page", style="opacity: .7; color: #000;")
```

