---
title: "14 Decision Errors"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(openintro)
knitr::opts_chunk$set(echo = FALSE)
```


## One-Sided Hypothesis Testing with Dr. Arbuthnot

### Exercise 1

Recall the Dr. Arbuthnot data set on birthrates. We are interested in testing whether boys are born at a higher rate than girls. Take a quick look at the `arbuthnot` data set in the console below:

```{r arbuthnot-glimpse, exercise=TRUE}

```

We are interested in answering the following research question:

> Are male babies more at a higher rate than female babies?

In the console below, answer the following:

1. What is the population of interest?
2. What is the sample we're using to answer this question?
3. How would you write this in a hypothesis testing framework?

```{r q-1, exercise=TRUE}
# 1:

# 2:

# 3:
```

### Exercise 2

In the console below,

1. Use the summarize function compute the test statistic for this test.
2. Assume that the population standard deviation is 0.250 and compute the standard error for our test statistic.

```{r q-2, exercise=TRUE}

```

In the console below,

3. Compute the p-value for this test.

```{r q-3, exercise=TRUE}

```

What is the outcome of our hypothesis test?

```{r q-4, exercise=TRUE}
# 
```

### Exercise 3

Use the console below to answer the following questions:

1. If the null were true, what would be the probability of accidentally rejecting the null?
2. What would be the cutoff for the critical region of this test?
3. What is the smallest proportion of boys that would result in rejecting the null hypothesis?
4. If the point estimate stayed the same, what would be the minimum sample size that would still lead to rejecting the null?

```{r q-5, exercise=TRUE}

```

And put your answers here:

```{r q-6, exercise=TRUE}
#
```

### Exercise 4

Perhaps a better research questions would have been:

> Are male and female babies born at the same rate?

What would the hypotheses be in this scenario?

```{r q-7, exercise=TRUE}
# 
```


### Exercise 4

Use the console below to find the critical regions for the hypothesis test you just wrote down above using a significance level fo $\alpha=0.01$:

```{r q-8, exercise=TRUE}

```

Find  a sample size for at which you would reject the null for the first (one-sided) test we did but would not reject the null for this two-sided test.

```{r q-9, exercise=TRUE}

```

### Exercise 5

Suppose you are Dr. Arbuthnot and have not yet collected your data. You think that the true proportion of babies that are male is 52%. You may assume a population standard deviation of 0.5. If this is the true proportion, what sample size would you need so that the power (i.e. the probability of rejecting the null) of your two-sided hypothesis test is 90%? Use the console below for any  calculations.

```{r calculator, exercise=TRUE}

```

Enter your answer here:

```{r answer, exercise=TRUE}
# 
```

## Which is higher? 

In each part below, there is a value of interest and two scenarios:  (i) and (ii). For each part, report if the value of interest is larger under scenario (i), scenario (ii), or whether the value is equal under the scenarios.

a. The standard error of $\hat{p}$ when (i) $n = 125$ or (ii) $n = 500$.

b. The margin of error of a confidence interval when the confidence level is (i) 90% or (ii) 80%.

c. The p-value for a Z-statistic of 2.5 calculated based on a (i) sample with $n = 500$ or based on a (ii) sample with $n = 1000$.

d. The probability of making a Type 2 Error when the alternative hypothesis is true and the significance level is (i) 0.05 or (ii) 0.10.
    
```{r hw-1, exercise=TRUE}
# a)

# b)

# c)

# d)
```

```{r hw-1-solution}
# a) Scenario (i) is higher. Recall that a sample mean based on less data tends to be less accurate and have larger standard errors.

# b) Scenario (i) is higher. The higher the confidence level, the higher the corresponding margin of error.

# c) They are equal. The sample size does not affect the calculation of the p-value for a given Z-score.

# d) Scenario (i) is higher. If the null hypothesis is harder to reject (lower α), then we are more likely to make a Type 2 Error when the alternative hypothesis is true.
```

## Testing for food safety.
A food safety inspector is called upon to investigate a restaurant with a few customer reports of poor sanitation practices. The food safety inspector uses a hypothesis testing framework to evaluate whether regulations are not being met. If he decides the restaurant is in gross violation, its license to serve food will be revoked.

a.  Write the hypotheses in words.

b.  What is a Type 1 Error in this context?

c.  What is a Type 2 Error in this context?

d.  Which error is more problematic for the restaurant owner? Why?

e.  Which error is more problematic for the diners? Why?

f.  As a diner, would you prefer that the food safety inspector requires strong evidence or very strong evidence of health concerns before revoking a restaurant's license? Explain your reasoning.

```{r hw-3, exercise=TRUE}
# a)

# b)

# c)

# d)

# e)

# f)
```

```{r hw-3-solution}
# a) H0: The restaurant meets food safety and sanitation regulations. HA: The restaurant does not meet food safety and sanitation regulations.
# 
# b) The food safety inspector concludes that the restaurant does not meet food safety and sanitation regulations and shuts down the restaurant when the restaurant is actually safe.
# 
# c) The food safety inspector concludes that the restaurant meets food safety and sanitation regulations and the restaurant stays open when the restaurant is actually not safe.
# 
# d) A Type 1 Error may be more problematic for the restaurant owner since his restaurant gets shut down even though it meets the food safety and sanitation regulations.
# 
# e) A Type 2 Error may be more problematic for diners since the restaurant deemed safe by the inspector is actually not.
# 
# f) A diner would prefer strong evidence as any indication of evidence might mean there may be an issue with the restaurant meeting food safety and sanitation regulations, and diners would rather a restaurant that meet the regulations get shut down than a restaurant that doesn’t meet the regulations not get shutdown.
```

## True or false.
Determine if the following statements are true or false, and explain your reasoning. If false, state how it could be corrected.

a. If a given value (for example, the null hypothesized value of a parameter) is within a 95% confidence interval, it will also be within a 99% confidence interval.

b. Decreasing the significance level ($\alpha$) will increase the probability of making a Type 1 Error.

c. Suppose the null hypothesis is $p = 0.5$ and we fail to reject $H_0$. Under this scenario, the true population proportion is 0.5.

d. With large sample sizes, even small differences between the null value and the observed point estimate, a difference often called the effect size, will be identified as statistically significant.
    
```{r hw-2, exercise=TRUE}
# a)

# b)

# c)

# d)
```

```{r hw-2-solution}
# a) True.
# 
# b) False. The significance level is the probability of the Type 1 Error.
# 
# c) False. Failure to reject H0 only means there wasn’t sufficient evidence to reject it, not that it has been confirmed.
# 
# d) True.
```

## Submit

```{css echo=FALSE}
@media print {
  .topicsContainer,
  .topicActions,
  .exerciseActions .skip {
    display: none;
  }
  .topics .tutorialTitle,
  .topics .section.level2,
  .topics .section.level3:not(.hide) {
    display: block;
  }
  .topics {
    width: 100%;
  }
  .tutorial-exercise, .tutorial-question {
    page-break-inside: avoid;
  }
  .section.level3.done h3 {
    padding-left: 0;
    background-image: none;
  }
  .topics .showSkip .exerciseActions::before {
    content: "Topic not yet completed...";
    font-style: italic;
  }
}
```

Make sure you have the tutorial open in a browser (preferably Google Chrome). Please press `print page` button below to print the tutorial. Print it to a `pdf` file and upload it into Moodle. Make sure all code boxed are visible. If you can't do that please submit your answers to [the Google Form](https://forms.gle/cDqdcVmkEaYAynfQ7).

```{js print2pdf1, context="server"}
    // the following 2 chunks print the completed sections of the tutorial to PDF
    // uses "css/print2pdf.css"
    $(document).on('shiny:inputchanged', function(event) {
      if (event.name === 'print2pdf') {
        window.print();
      }
    });
```
    
```{r print2pdf2}
    # button can be placed anywhere in the tutorial
    actionButton("print2pdf", "Print page", style="opacity: .7; color: #000;")
```
    