---
title: "Probability"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(openintro)
library(tidyverse)
library(janitor)

knitr::opts_chunk$set(echo = FALSE)
```


## Smallpox

The smallpox data set provides a sample of 6,224 individuals from the year 1721 who were exposed to smallpox in Boston. Doctors at the time believed that inoculation, which involves exposing a person to the disease in a controlled form, could reduce the likelihood of death. In theconsole below, take a minute to explore the ``smallpox`` dataset:

```{r explore, exercise=TRUE}

```

In the console below, generate a contingency table with `inoculated` in the columns and `result` in the rows. The `janitor` package has been loaded so feel free to use `adorn_totals` if you'd like:

```{r cont-table, exercise=TRUE}

```

```{r cont-table-solution}
smallpox %>% 
  count(result, inoculated) %>% 
  pivot_wider(id_cols = result, names_from = inoculated, values_from = n) %>% 
  adorn_totals(where = c("row", "col"))
```

On a separate sheet of paper, compute the table proportions for this data.

Now use R to compute the table proportions to verify your solutions:

```{r table-props, exercise=TRUE}

```

```{r table-props-solution}
smallpox %>% 
  count(result, inoculated) %>% 
  mutate(prop = n/sum(n)) %>% 
  pivot_wider(id_cols = result, names_from = inoculated, values_from = prop) %>% 
  adorn_totals(where = c("row", "col"))
```

```{r questions-1}
quiz(
  question_numeric("What is the probability that a randomly selected individual died of small pox?",
                 answer(0.1365681, correct=TRUE),
                 allow_retry = TRUE,
                 tolerance = 0.001),
  question("What type of probability is the above probability?",
           answer("Marginal", correct=TRUE),
           answer("Joint"),
           answer("Conditional"),
           allow_retry=TRUE),
  
  question_numeric("What is the probability that a randomly selected individual was inoculated and lived?",
                 answer(0.0382390746, correct=TRUE),
                 allow_retry = TRUE,
                 tolerance = 0.001),
  question("What type of probability is the above probability?",
           answer("Marginal"),
           answer("Joint", correct=TRUE),
           answer("Conditional"),
           allow_retry=TRUE),
  
  question_numeric("What is the probability that a randomly selected individual who lived was inoculated?",
                 answer(0.044287309, correct=TRUE),
                 allow_retry = TRUE,
                 tolerance = 0.001),
  question("What type of probability is the above probability?",
           answer("Marginal"),
           answer("Joint"),
           answer("Conditional", correct=TRUE),
           allow_retry=TRUE),
  
  question_numeric("What is the probability that a randomly selected individual was not inoculated and lived?",
                 answer(0.8251928, correct=TRUE),
                 allow_retry = TRUE,
                 tolerance = 0.001),
  question("What type of probability is the above probability?",
           answer("Marginal"),
           answer("Joint", correct=TRUE),
           answer("Conditional"),
           allow_retry=TRUE),
  
  question_numeric("What is the probability that a randomly selected individual who wasn’t inoculated lived?",
                 answer(0.8588629, correct=TRUE),
                 allow_retry = TRUE,
                 tolerance = 0.001),
  question("What type of probability is the above probability?",
           answer("Marginal"),
           answer("Joint"),
           answer("Conditional", correct=TRUE),
           allow_retry=TRUE)
)
```

Was it better to be inoculated or not inoculated? Make an argument using marginal, joint, or conditional probabilities and be prepared to discuss with the class.
