---
title: "Confidence Intervals"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(openintro)
library(tidyverse)
library(infer)

healthcare_law_survey <-healthcare_law_survey %>% 
  mutate(response = ifelse(response == "approve", "approve", "don't approve"))

knitr::opts_chunk$set(echo = FALSE)
```


## Pew Research Center poll on health care

We will be analyzing the data set `healthcare_law_survey`. Metadata for this data set can be found [here](https://www.openintro.org/data/index.php?data=healthcare_law_survey). We would like to estimate the proportion of Americans (in 2012) who approve of the ACA policy of requiring all Americans to have health insurance or face a fine. Please use the console below to take a look at the data set:

```{r sandbox, exercise=TRUE}

```

What is the parameter of interest and what is the sample in this scenario:

```{r parameter-sample, exercise=TRUE}
# Parameter:

# Sample:
```

Let's call the proportion of all Americans that approve of the policy $p_{app}$ and our point estimate of this quantity $\hat{p}_{app}$. In the console below, compute $\hat{p}_{app}$ (you may want to look at my slides from this lesson):

```{r p-hat, exercise=TRUE}

```

What is our point estimate and a 95% confidence interval for the $p_{app}$:

```{r point-estimate-ci, exercise=TRUE}
# Point estimate:

# 95% Confidence interval:
```

Use the console below to compute a 99% confidence interval:

```{r ci-99, exercise=TRUE}

```










