---
title: "20: Inference For Comparing Two Independent Means"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(openintro)

chickwts <- chickwts %>% 
  filter(feed == "horsebean" | feed == "linseed")

set.seed(1234)
epa2021_sample <- epa2021 %>%
  filter(transmission_desc %in% c("Manual", "Automatic")) %>%
  group_by(transmission_desc) %>%
  sample_n(size = 25)


knitr::opts_chunk$set(echo = FALSE)
```


## Chicken diet: horsebean vs. linseed. 

Chicken farming is a multi-billion dollar industry, and any methods that increase the growth rate of young chicks can reduce consumer costs while increasing company profits, possibly by millions of dollars. An experiment was conducted to measure and compare the effectiveness of various feed supplements on the growth rate of chickens. Newly hatched chicks were randomly allocated into six groups, and each group was given a different feed supplement. In this exercise we consider chicks that were fed horsebean and linseed. The dataset can be found in the data frame `chickwts`. Take a look below:


```{r sandbox-2, exercise=TRUE}

```

Your task is to test whether there is any difference between the weights of chicken under a horsebean vs linseed diet. If you feel comfortable doing this on you own please do so in the console below. Don't forget to check whether the data meets the normality assumptions:

```{r self-2, exercise=TRUE}

```

In the console below or on a separate sheet of paper write down the null and alternative hypotheses for this test:

```{r hyp, exercise=TRUE}
#
```

```{r hyp-solution}
# H0: mu_lin - mu_horse = 0

# HA: mu_lin - mu_horse ≠ 0
```


In the console below, create a histograms (use 5 bins) and box plots of both groups:

```{r ben-viz, exercise=TRUE}

```

```{r ben-viz-hint-1}
ggplot(_____) +
  _____(_____)

ggplot(_____) +
  _____(_____)
```

```{r ben-viz-hint-2}
ggplot(_____, _____) +
  geom_histogram(_____)

ggplot(_____, _____) +
  geom_boxplot()
```

```{r ben-viz-hint-3}
ggplot(_____, aes(x=_____, fill=______)) +
  geom_histogram(bins=_____, alpha=0.5, position="identity")

ggplot(_____, aes(x=_____, y=______)) +
  geom_boxplot()
```

```{r ben-viz-hint-4}
ggplot(chickwts, aes(x=weight, fill=feed)) +
  geom_histogram(bins=6, alpha=0.5, position="identity")

ggplot(chickwts, aes(x=weight, y=feed)) +
  geom_boxplot()
```

While the data set is small, it does not seems to be skewed and there are no outliers so we should feel comfortable applying the central limit theorem.

Recall that the formula for the t-statistic is \[T = \frac{(\bar{x}_1 - \bar{x}_2) - 0}{\sqrt{s_1^2/n_1 + s_2^2/n_2}}.\] In the console below compute the t-statistic of the data.

```{r t-stat, exercise=TRUE}

```

```{r t-stat-hint-1}
chickwts %>% 
  _____ %>% 
  _____


T = ______
T
```

```{r t-stat-hint-2}
chickwts %>% 
  group_by(_____) %>% 
  summarize(______, _____, _____)


T = ______
T
```

```{r t-stat-hint-3}
chickwts %>% 
  group_by(feed) %>% 
  summarize(mean(weight), var(weight), n())


T = ______
T
```

```{r t-stat-hint-4}
chickwts %>% 
  group_by(feed) %>% 
  summarize(mean(weight), var(weight), n())


T = (218.75 - 160.20) / sqrt(2728.568 / 12		+ 1491.956 / 10)
T
```

```{r df-q}
question_numeric("When computing our p-value, how many degrees of freedom should we use?",
                 answer(9, correct=TRUE),
                 allow_retry = TRUE)
```

In the console below, compute the p-value for this test. On a separate sheet of paper sketch the t-distribution and shade the region corresponding to the p-value:

```{r p-val-setup}
T = (218.75 - 160.20) / sqrt(2728.568 / 12		+ 1491.956 / 10)
````

```{r p-val, exercise=TRUE}

```

```{r p-val-hint-1}
2*pt(_____, df=_____)
```

```{r p-val-hint-2}
2*pt(-T, df=9)

2* (1-pt(T, df=9)) # Also works
```

```{r outcome-1}
question_radio("What would be the outcome of the test with a significance level of 0.05?",
               answer("Fail to reject"),
               answer("Reject the null", correct=TRUE),
               allow_retry = TRUE)
```

```{r outcome-2}
question_radio("What would be the outcome of the test with a significance level of 0.01?",
               answer("Fail to reject", correct=TRUE),
               answer("Reject the null"),
               allow_retry = TRUE)
```

## Fuel efficiency in the city.

Each year the US Environmental Protection Agency (EPA) releases fuel economy data on cars manufactured in that year. A random sample of cars with manual and automatic transmissions manufactured in 2021 is contained in the data frame `epa2021_sample`. We're interested in whether there is a difference in fuel efficiency in the city between manual and automatic cars. The variables we're interested in are `transmission` and `city_mpg`. Take a peak below:

```{r sandbox-1, exercise=TRUE}

```


If you feel up to it, please create the 95% confidence interval in the console below for the difference in city mileage between cars with automatic and manual transmissions. This includes doing a little exploratory data analysis and verifying that the data is meets the normality conditions. If you don't feel comfortable doing this the remaining consoles will walk you through the steps.

```{r self-1, exercise=TRUE}

```

In the console below generate histograms and box plots of the `city_mpg` variable for both automatic and manual transmissions. For the histogram, use 5 bins:

```{r stage-viz, exercise=TRUE}

```

```{r stage-viz-hint-1}
ggplot(_____) +
  _____(_____)

ggplot(_____) +
  _____(_____)
```

```{r stage-viz-hint-2}
ggplot(_____, _____) +
  geom_histogram(_____)

ggplot(_____, _____) +
  geom_boxplot()
```

```{r stage-viz-hint-3}
ggplot(_____, aes(x=_____, fill=_____)) +
  geom_histogram(bins=_____, alpha=0.5, position="identity")

ggplot(_____, aes(x=_____, y=_____)) +
  geom_boxplot()
```

```{r stage-viz-hint-4}
ggplot(epa2021_sample, aes(x=city_mpg, fill=transmission)) +
  geom_histogram(bins=5, alpha=0.5, position="identity")

ggplot(epa2021_sample, aes(x=city_mpg, y=transmission)) +
  geom_boxplot()
```

```{r q-1}
question_radio("(Practice Test Question): Does this data seem to meet the normality conditions? On a separate sheet of paper explain your answer.",
               answer("Yes", correct=TRUE),
               answer("No"),
               allow_retry = TRUE)
```

The data has little to no skew but there are some outliers in the manual group. In addition, the sample size is 25 for each group which is borderline. Since none of the outliers are extreme and both histograms look relatively symmetric we should feel comfortable assuming the data is normal. However this is a bit of a borderline case.

Recall the the formula for a confidence interval is $\bar{x}_{man}-\bar{x}_{aut} \pm t^* SE$ where $SE = \sqrt{s_1^2/n_1 + s_2^2/n_2}$. In the console below, compute $\bar{x}_{man}-\bar{x}_{aut}$ and $SE$:

```{r xbar-se-1, exercise=TRUE}

```

```{r xbar-se-1-hint-1}
epa2021_sample %>% 
  _____

point_est <- _____
SE <- _____

point_est
SE
```

```{r xbar-se-1-hint-2}
epa2021_sample %>% 
  group_by(_____) %>% 
  summarize(______)

point_est <- _____
SE <- _____

point_est
SE
```

```{r xbar-se-1-hint-3}
epa2021_sample %>% 
  group_by(_____) %>% 
  summarize(______, _____, _____)

point_est <- _____
SE <- _____

point_est
SE
```

```{r xbar-se-1-hint-4}
epa2021_sample %>% 
  group_by(transmission) %>% 
  summarize(mean(city_mpg), var(city_mpg), n())

point_est <- _____
SE <- _____

point_est
SE
```

```{r xbar-se-1-hint-5}
epa2021_sample %>% 
  group_by(transmission) %>% 
  summarize(mean(city_mpg), var(city_mpg), n())

point_est <- 22.7-17.4
SE <- sqrt(21.0/25 + 11.8/25)

point_est
SE
```


Now let's find $t^\star$ for a 95% confidence interval.

```{r q-2}
question_numeric("How many degrees of free would you use in this t-distribution?",
                 answer(24, correct=TRUE),
                 allow_retry = TRUE)
```

In the console below, find $t^\star$:

```{r t-1, exercise=TRUE}

```

```{r t-1-hint-1}
tstar <- qt(_____, df=______)
tstar
```

```{r t-1-hint-2}
tstar <- qt(.975, df=24)
tstar
```

In the console below, compute the final confidence interval:

```{r ci-1-setup}
point_est <- 22.7-17.4
SE <- sqrt(21.0/25 + 11.8/25)
tstar <- qt(.975, df=24)
```

```{r ci-1, exercise=TRUE}

```


```{r ci-1-hint-1}
_____ - _____ * _____

_____ + _____ * _____
```

```{r ci-1-hint-2}
point_est - tstar * SE

point_est + tstar * SE
```

So our point estimate for the average increase in city MPG in manual cars is 5.3 with a 95% confidence intervals of 2.94 to 7.66 MPG.


## Diamonds, mathematical interval.
We have data on two random samples of diamonds: one with diamonds that weigh 0.99 carats and one with diamonds that weigh 1 carat. 
Each sample has 23 diamonds. 
Sample statistics for the price per carat of diamonds in each sample are provided below.
Assuming that the conditions for conducting inference using a mathematical model are satisfied, construct a 95% confidence interval for the true population difference in prices per carat of diamonds that weigh 0.99 carats and 1 carat. (Wickham 2015)

```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(scales))
suppressMessages(library(kableExtra))

set.seed(47)
diamonds_100_99_sample <- diamonds %>%
  filter(carat %in% c(1, 0.99)) %>%
  group_by(carat) %>%
  slice_sample(n = 23) %>% 
  mutate(
    ppc   = price/carat/100,
    carat = if_else(carat == 1, paste(carat, "carat"), paste(carat, "carats"))
    )

diamonds_100_99_summary <- diamonds_100_99_sample %>%
  group_by(carat) %>%
  summarise(
    Mean = format(mean(ppc), digits = 4, nsmall = 2),
    SD = format(sd(ppc), digits = 4, nsmall = 2),
    n = n()
  ) %>%
  mutate(
    Mean = paste0("$", Mean),
    SD = paste0("$", SD)
    )
    
diamonds_100_99_summary %>%
  kbl(linesep = "", booktabs = TRUE, align = "lccc", col.names = c("", "Mean", "SD", "n"), digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), 
                latex_options = "HOLD_position",
                full_width = FALSE) %>%
  column_spec(1:4, width = "6em")
```

```{r q1, exercise=TRUE}

```

## Diamonds solution

The 95% confidence interval can be calculated as follows:

\[
\begin{aligned}
t^{\star}_{df} &= t^{\star}_{23 - 1 = 22} = 2.07 \\
(\bar{x}_{0.99} - \bar{x}_{1}) \pm t^{\star}_{df} \sqrt{ \frac{s_{0.99}^2}{n_{0.99} } + \frac{s_{1}^2}{n_{1}} } &= (44.51 - 57.2) \pm 2.07 \sqrt{ \frac{177.4224}{23} + \frac{330.8761}{23} } \\
&= -12.69 \pm 2.07 \times 4.7011 \\
&= -12.69 \pm 9.73 \\
&= (-22.42, -2.96)
\end{aligned}
\]

We are 95% confident that the population average price per carat of 0.99 carat diamonds is $-22.42 to $-2.96 lower than the population average price per carat of 1 carat diamonds.



## True / False: comparing means.
Determine if the following statements are true or false, and explain your reasoning for statements you identify as false.

a.  As the degrees of freedom increases, the $t$-distribution approaches normality.

b.  If a 95% confidence interval for the difference between two population means contains 0, a 99% confidence interval calculated based on the same two samples will also contain 0.

c.  If a 95% confidence interval for the difference between two population means contains 0, a 90% confidence interval calculated based on the same two samples will also contain 0.

```{r q2, exercise=TRUE}
# a)

# b)

# c)
```

```{r q2-solution}
# a) True

# b) True

# c) False. 90% confidence interval is narrower, therefore it may or may not include 0.
```
    

## Gaming, distracted eating, and recall.
A group of researchers who are interested in the possible effects of distracting stimuli during eating, such as an increase or decrease in the amount of food consumption, monitored food intake for a group of 44 patients who were randomized into two equal groups. 
The 22 patients in the treatment group who ate their lunch while playing solitaire were asked to do a serial-order recall of the food lunch items they ate. 
The average number of items recalled by the patients in this group was 4. 9, with a standard deviation of 1.8. 
The average number of items recalled by the patients in the control group (no distraction) was 6.1, with a standard deviation of 1.8. 
Do these data provide strong evidence that the average numbers of food items recalled by the patients in the treatment and control groups are different?
Assume that conditions for conducting inference using mathematical models are satisfied. (Oldham-Cooper et al. 2011)

```{r q4, exercise=TRUE}

```

## Gaming, distracted eating, and recall. Solutions
The hypotheses are $H_0: \mu_{T} = \mu_{C}$ and $H_A: \mu_{T} \ne \mu_{C}$. We are told to assume that conditions for inference are satisfied.

\[
\begin{aligned}
T &= \frac{(\bar{x}_T - \bar{x}_C) - (\mu_T - \mu_C)}{\sqrt{ \frac{s_T^2}{n_T} + \frac{s_C^2}{n_C} }} \\
&= \frac{(4.9 - 6.1) - 0}{ \sqrt{\frac{1.8 ^2}{22} + \frac{1.8 ^2}{22}} } = \frac{-1.2}{0.543} = -2.21 \\
df &= min(n_1 - 1, n_2 - 1) = min(22 - 1, 22 - 1) = 21 \\
p-value &= P(|T_{21}| > 2.21) \rightarrow 0.02 < p-value < 0.05
\end{aligned}
\]

Since p-value < 0.05, we reject $H_0$. The data provide convincing evidence that the average numbers of food items recalled by the patients in the treatment and control groups are different.

## Submit

```{css echo=FALSE}
@media print {
  .topicsContainer,
  .topicActions,
  .exerciseActions .skip {
    display: none;
  }
  .topics .tutorialTitle,
  .topics .section.level2,
  .topics .section.level3:not(.hide) {
    display: block;
  }
  .topics {
    width: 100%;
  }
  .tutorial-exercise, .tutorial-question {
    page-break-inside: avoid;
  }
  .section.level3.done h3 {
    padding-left: 0;
    background-image: none;
  }
  .topics .showSkip .exerciseActions::before {
    content: "Topic not yet completed...";
    font-style: italic;
  }
}
```

Make sure you have the tutorial open in a browser (preferably Google Chrome). Please press `print page` button below to print the tutorial. Print it to a `pdf` file and upload it into Moodle. Make sure all code boxed are visible. If you can't do that please submit your answers to [the Google Form](https://forms.gle/nsxVUsD9SKHMQVrcA).

```{js print2pdf1, context="server"}
    // the following 2 chunks print the completed sections of the tutorial to PDF
    // uses "css/print2pdf.css"
    $(document).on('shiny:inputchanged', function(event) {
      if (event.name === 'print2pdf') {
        window.print();
      }
    });
```
    
```{r print2pdf2}
    # button can be placed anywhere in the tutorial
    actionButton("print2pdf", "Print page", style="opacity: .7; color: #000;")
```