---
title: "Inference For A Single Proportion"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(openintro)
library(infer)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)

mice <- tibble(mouse_id = seq(1, 127), result = sample(c(rep("perfume", 80), rep("no perfume", 127-80))))

global_warming_pew <- global_warming_pew %>% 
  mutate(response_binary = ifelse(response == "Earth is warming", "Earth is warming", "Not warming / Don't know / refuse to answer "))

# cwd <- read_csv("PRNPII-data.csv")
# 
# cwd_wi <- cwd %>% 
#   filter(STATE == "WI") %>% 
#   mutate(G96S = ifelse(G96S == "A", "A", "G"))
# 
# cwd <- cwd %>% 
#   filter(Status != "NOT TESTED") %>% 
#   mutate(Q95H = ifelse(Q95H == "A", "A", "M"),
#          G96S = ifelse(G96S == "A", "A", "G"))
# 
# cwd_A <- cwd %>% 
#   filter(G96S == "A")
# 
# obs_prop <- cwd_A %>% 
#   specify(response = Status, success = "POSITIVE") %>% 
#   calculate(stat = "prop")
```


## Olfactory Responses in Mice

Suppose you're a researcher who is researching the olfactory centers of mice. You are interested in investigating whether you can attract mice to a certain bowl of food by spraying the food with a "perfume" that contains pheromones from other mice. In your experiment a mouse is placed in between two bowls of food, one with the perfume and one without. The bowl that the mouse eats out of first is then recorded. Each mouse is independent of all other mice. The results of your experiment are stored in the data frame `mice`. Take a look at the data set below. 

```{r sandbox, exercise=TRUE}

```

Your task is to conduct a hypothesis test for your research question. If you feel up to it, try and do it all in the console below. This includes selecting your own significance level. Try and do this both with and without the `infer` package. If you don't feel comfortable doing this the remaining consoles will walk you through the steps.

```{r advanced, exercise=TRUE}

```


Call the proportion of mice (in the population) that prefer the food with the perfume $p$. On a sheet of paper or in the console, write down what you think your null and alternative hypotheses are:

```{r hypotheses, exercise=TRUE}
# H0:
# HA:
```


```{r hypotheses-solution}
# H0: p=.5
# HA: p>.5
```

Let's also assume $\alpha = 0.05$. We'll first do every without using the `infer` package. In the console below, compute the test statistic:

```{r test-statistic, exercise=TRUE}

```

```{r test-statistic-hint-1}
_____ %>% 
  _____
```

```{r test-statistic-hint-2}
_____ %>% 
  _____(point_estimate = mean(____ == "_____"))
```

```{r test-statistic-hint-3}
mice %>% 
  summarize(point_estimate = mean(result == "perfume"))
```

In the console below, compute the standard error for this test:

```{r standard-error, exercise=TRUE}

```

```{r standard-error-hint-1}
sqrt(____ * (1 - _____)/_____)
```

```{r standard-error-hint-2}
sqrt(.5 * (1 - .5)/127)
```

Does the CLT apply in this scenario? 

```{r clt, exercise=TRUE}
#
```

```{r clt-solution}
# Yes, the success failure condition is easily statisfied because n= 127 
# and p = 0.5 so np >=10 and n(1-p) >= 10

```

Now compute your z-score and use the `normTail` function to display the region of the standard normal distribution corresponding to your p-value:

```{r z, exercise=TRUE}

```

```{r z-hint-1}
z_score <- 
z_score
normTail(____, ____, _____)
```

```{r z-hint-2}
z_score <- (_____ - _____)/_____
z_score
normTail(m=_____, s=____, U=_____)
```

```{r z-hint-3}
z_score <- (_____ - 0.5)/0.04436783
z_score
normTail(m=0, s=1, U=_____)
```

```{r z-hint-4}
z_score <- (0.630 - 0.5)/0.04436783
z_score
normTail(m=0, s=1, U=z_score)
```

Compute your p-value:

```{r p-val, exercise=TRUE}

```

```{r p-val-hint-1}
pnorm(____, ____, ____)
```

```{r p-val-hint-2}
pnorm(____, mean = ____, sd = ____)
```

```{r p-val-hint-3}
1-pnorm(2.930051, mean = 0, sd = 1)
```

```{r q-1}
question_radio(
"Would is the result of this hypothesis test with a significance level of $\\alpha = 0.05$",
answer("Fail to Reject the Null Hypothesis"),
answer("Reject the null hypothesis", correct=TRUE),
allow_retry=TRUE
)
```

Now we'll repeat the process using the `infer` package. In the console below:

1. Define the null distribution.
2. (Optional) Visualize the null distribution.
3. Compute the test statistic.
4. (Optional) Visualize the p-value.
5. Compute the p_value.

```{r infer-1, exercise=TRUE}

```

```{r infer-1-hint-1}
null_dist <- mice %>% 
  specify_____, _____) %>% 
  hypothesize(_____, _____) %>% 
  assume(_____)

null_dist %>% 
  visualize()

test_stat <- mice %>% 
  specify_____, _____) %>% 
  hypothesize(______, ______) %>% 
  calculate(_____)

null_dist %>% 
  visualize() +
    shade_p_value(______, ______)

null_dist %>% 
  get_p_value(_____, _____)
```

```{r infer-1-hint-2}
null_dist <- mice %>% 
  specify(response = _____, success = "_____") %>% 
  hypothesize(null = "_____", p = _____) %>% 
  assume(distribution = "_____")

null_dist %>% 
  visualize()

test_stat <- mice %>% 
  specify(response = ______, success = "_____") %>% 
  hypothesize(null = "_____", p = _____) %>% 
  calculate(stat = "_____")

null_dist %>% 
  visualize() +
    shade_p_value(obs_stat = _____, direction = "______")

null_dist %>% 
  get_p_value(obs_stat = ______, direction = "______")
```

```{r infer-1-hint-3}
null_dist <- mice %>% 
  specify(response = result, success = "perfume") %>% 
  hypothesize(null = "point", p = 0.5) %>% 
  assume(distribution = "z")

null_dist %>% 
  visualize()

test_stat <- mice %>% 
  specify(response = result, success = "perfume") %>% 
  hypothesize(null = "point", p = 0.5) %>% 
  calculate(stat = "z")

null_dist %>% 
  visualize() +
    shade_p_value(obs_stat = test_stat, direction = "right")

null_dist %>% 
  get_p_value(obs_stat = test_stat, direction = "right")
```

## Global warming poll

A 2010 Pew Research poll asked 1,306 Americans, "From what you've read and heard, is there solid evidence that the average temperature on earth has been getting warmer over the past few decades, or not?" The resulting data can be found in the data frame `global_warming_pew`. Take a look at the data below:

```{r sandbox-gw, exercise=TRUE}

```

Your task is construct a 90% confidence interval for the proportion of Americans (in 2010) who believe that the earth is warming. If you feel up to it, try and do it all in the console below. Try and do this both with and without the `infer` package. If you don't feel comfortable doing this the remaining consoles will walk you through the steps.

```{r ci-advanced, exercise=TRUE}

```

Well first construct a confidence interval without using the `infer` package. Remember that confidence intervals take the form *point estimate* $\pm z^*\times SE$. Let's first compute $\hat{p}$, our point estimate, which will be the proportion of responders in our data set that say the "Earth is warming". Compute this number in the console below:

```{r point_est, exercise=TRUE}

```

```{r point_est-hint-1}
_____ %>% 
  _____
```

```{r point_est-hint-2}
_____ %>% 
  summarize(mean(_____=="_____"))
```

```{r point_est-hint-3}
global_warming_pew %>% 
  summarize(mean(response=="Earth is warming"))
```

Now that we have our point estimate, compute the standard error (technically the estimate of the standard error):

```{r standard-err, exercise=TRUE}

```

```{r standard-err-hint-1}
sqrt(_____*(1-_____)/_____)
```

```{r standard-err-hint-2}
sqrt(0.5996449*(1-0.5996449)/2253)
```

Compute $z^*$ for a 90% confidence interval:

```{r z-star, exercise=TRUE}

```

```{r z-star-hint-1}
qnorm(_____, _____, _____)
```

```{r z-star-hint-2}
qnorm(_____, mean=_____, sd=_____)
```

```{r z-star-hint-3}
qnorm(.95, mean=0, sd=1)
```

Now that we have all the ingredients that we need, compute the confidence interval:

```{r con-int, exercise=TRUE}

```

```{r con-int-hint-1}
_____ - _____ * _____
_____ + _____ * _____
```

```{r con-int-hint-2}
0.5996449 - 1.644854 * 0.0103226
0.5996449 + 1.644854 * 0.0103226
```


So our 90% confidence interval is (58.3%, 61.7%).

Now we'll repeat the process using the `infer` package. In what follows, use the `response_binary` column rather than the `response` column. In the console below:

1. Define the sampling distribution. 
2. Compute the point estimate.
3. Compute the confidence interval.
4. (Optional) Visualize sampling distribution and confidence interval.

```{r infer-ci, exercise=TRUE}

```

```{r infer-ci-hint-1}
samp_dist <- global_warming_pew %>% 
  # specify the variable you're interested in
  specify(_____, _____) %>% 
  # specify sampling distribution
  assume(______)

point_est <- global_warming_pew %>% 
  # specify the variable you're interested in
  specify(______, _____) %>% 
  # specify the statistic you want to compute
  calculate(_____)

point_est

ci <- samp_dist %>% 
  get_ci(_____, _____)

ci

samp_dist %>% 
  visualize() +
    shade_confidence_interval(_____) 
```

```{r infer-ci-hint-2}
samp_dist <- global_warming_pew %>% 
  # specify the variable you're interested in
  specify(response = _____, success = "______") %>% 
  # specify sampling distribution
  assume(distribution = "_____")

point_est <- global_warming_pew %>% 
  # specify the variable you're interested in
  specify(response = _____, success = "_____") %>% 
  # specify the statistic you want to compute
  calculate("______")

point_est

ci <- samp_dist %>% 
  get_ci(level = _____, point_estimate = _____)

ci

samp_dist %>% 
  visualize() +
    shade_confidence_interval(_____) 
```

```{r infer-ci-hint-3}
samp_dist <- global_warming_pew %>% 
  # specify the variable you're interested in
  specify(response = response_binary, success = "Earth is warming") %>% 
  # specify sampling distribution
  assume(distribution = "z")

point_est <- global_warming_pew %>% 
  # specify the variable you're interested in
  specify(response = response_binary, success = "Earth is warming") %>% 
  # specify the statistic you want to compute
  calculate("prop")

point_est

ci <- samp_dist %>% 
  get_ci(level = 0.9, point_estimate = point_est)

ci

samp_dist %>% 
  visualize() +
    shade_confidence_interval(ci) 
```

<!-- ## Estimating Prevelance of G96S Mutations (Special Topic: Optional) -->

<!-- You are doing research with Dr. Friedlander's esteemed colleague, Professor Adam Brandt, investigating Chronic Wasting Disease (CWD) in Wisconsin deer. A newly discovered mutation at the G96S SNP (having an `A` instead of a `G` at that site) shows promise in reducing the susceptibility of deer to CWD. Dr. Brandt assigns you to estimate the prevalence of this mutation in the deer population in Wisconsin. You can use the `cwd_wi` data frame which stores information on this mutation in the column `G96S`. Take a look around the dataset below: -->

<!-- ```{r sandbox-3, exercise=TRUE} -->

<!-- ``` -->

<!-- After discussing the matter with Dr. Brandt you decide to construct a 95% confidence interval for the prevalence of this `A` mutation at the G96S position in the deer genome. In the console below, determine whether this data meets the *success-failure* condition (you may assume that the observations are independent): -->

<!-- ```{r clt-1, exercise=TRUE} -->

<!-- ``` -->

<!-- ```{r clt-1-hint-1} -->
<!-- cwd_wi %>%  -->
<!--   count(G96S) -->
<!-- ``` -->

<!-- ```{r q-2} -->
<!-- question_radio( -->
<!-- "Can you apply the CLT to this data", -->
<!-- answer("Yes, the independence and success-failure conditions are both met"), -->
<!-- answer("No, the independence condition is not met"), -->
<!-- answer("No, the success-failure condition is not met", correct=TRUE), -->
<!-- answer("No, neither the independence or success-failure conditions are met"), -->
<!-- allow_retry=TRUE -->
<!-- ) -->
<!-- ``` -->

<!-- In the console below, if you feel confident (PUN INTENDED!!!!), construct a 95% bootstrapped confidence interval for the prevalence of this mutation. If you don't yet feel comfortable, walk through the steps below: -->

<!-- ```{r advanced-boot-ci, exercise=TRUE} -->

<!-- ``` -->

<!-- First, compute a point estimate for the proportion of deer with this mutation: -->

<!-- ```{r point-est-boot, exercise=TRUE} -->

<!-- ``` -->

<!-- ```{r point-est-boot-hint-1} -->
<!-- cwd_wi %>%  -->
<!--   summarize(_____) -->
<!-- ``` -->

<!-- ```{r point-est-boot-hint-2} -->
<!-- cwd_wi %>%  -->
<!--   summarize(mean(G96S == "A")) -->
<!-- ``` -->

<!-- Use the infer package to generate the 10,000 bootstrapped replicates of the data to approximate the sampling distribution for the prevalence of this mutation (call the resulting data frame `sampling_dist`) and then visualize this distribution: -->

<!-- ```{r sampling-dist, exercise=TRUE} -->

<!-- ``` -->

<!-- ```{r sampling-dist-hint-1} -->
<!-- sampling_dist <- _____ %>%  -->
<!--   specify(response = _____, success= "_____") %>%  -->
<!--   calculate(stat = "_____") -->

<!-- sampling_dist %>%  -->
<!--   _____ -->
<!-- ``` -->

<!-- ```{r sampling-dist-hint-2} -->
<!-- sampling_dist <- cwd_wi %>%  -->
<!--   specify(response = G96S, success = "A") %>%  -->
<!--   generate(reps = 10000, type = "bootstrap") %>%  -->
<!--   calculate(stat = "prop") -->

<!-- sampling_dist %>%  -->
<!--   visualize(bins=100) -->
<!-- ``` -->

<!-- Now use the infer package to compute a 95% confidence interval based on this bootstrapped distribution and plot the shaded region using the `shade_ci` function: -->

<!-- ```{r ci-boot-setup} -->
<!-- sampling_dist <- cwd_wi %>%  -->
<!--   specify(response = G96S, success = "A") %>%  -->
<!--   generate(reps = 10000, type = "bootstrap") %>%  -->
<!--   calculate(stat = "prop") -->
<!-- ``` -->

<!-- ```{r ci-boot, exercise=TRUE} -->

<!-- ``` -->

<!-- ```{r ci-boot-hint-1} -->
<!-- ci <- _____ %>%  -->
<!--   _____(_____ = ____) -->
<!-- ci -->

<!-- _____ %>%  -->
<!--  _____ +   -->
<!--   _____(_____) -->
<!-- ``` -->

<!-- ```{r ci-boot-hint-2} -->
<!-- ci <- _____ %>%  -->
<!--   get_ci(_____ = _____) -->
<!-- ci -->

<!-- sampling_dist %>%  -->
<!--   visualize(bins=100) +   -->
<!--   shade_ci(_____) -->
<!-- ``` -->

<!-- ```{r ci-boot-hint-3} -->
<!-- ci <- sampling_dist %>%  -->
<!--   get_ci(level = 0.95) -->
<!-- ci -->

<!-- sampling_dist %>%  -->
<!--   visualize(bins=100) +   -->
<!--   shade_ci(ci) -->
<!-- ``` -->

<!-- What is our point estimate and 95% confidence interval: -->

<!-- ```{r boot-ci-final, exercise=TRUE} -->
<!-- #  -->
<!-- ``` -->

<!-- ```{r boot-ci-final-solution} -->
<!-- # Our point estimate is 3.03% with a 95% confidence interval of 0.76% to 6.06%.  -->
<!-- ``` -->


 
