---
title: "23 Using R for Inference"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(tidyverse)
library(openintro)
library(infer)

gerrymander <- read_csv("https://www.openintro.org/data/csv/gerrymander.csv")

helium <- helium %>%  
  pivot_longer(!trial, names_to = "type", values_to = "distance") %>% 
  select(!trial)


knitr::opts_chunk$set(echo = FALSE)
```


## Age at marriage

The dataset `age_at_mar` contains the age at first marriage of 5,534 US women who responded to the National Survey of Family Growth (NSFG) conducted by the CDC in the 2006 and 2010 cycle. You're interested in the age at which U.S. women get married. What types of analyses might you use to investigate this topic?

```{r type-age, exercise=TRUE}
# 
```

In the console below conduct such an analysis:

```{r analy-age, exercise=TRUE}

```


## Drug use of students and parents

The dataset `drug_use` contains information on 445 student parent pairs. You are interested in investigating the relationship between drug use in students and drug use in their parents. Come up with a research question appropriate for this data and then conduct a statistical analysis based on this question:

```{r sp-drugs, exercise=TRUE}

```

## Helium Football

At the 1976 Pro Bowl, Ray Guy, a punter for the Oakland Raiders, punted a ball that hung in mid-air long enough for officials to question whether the pigskin was filled with helium. The ball was found to be filled with air, but since then many have tossed around the idea that a helium-filled football would outdistance an air-filled one. Students at Ohio State University conducted an experiment to test this myth. They used two identical footballs, one air filled with air and one filled with helium. Each football was kicked 39 times and the two footballs were alternated with each kick. The data from this experiment is stored in `helium`. Perform and analysis below to determine if a ball filled with helium would travel further:

```{r helium, exercise=TRUE}

```

## Choose your own

Choose a dataset from [here](https://www.openintro.org/data/) that interests you, come up with a research question, and use statistics to analyze the questions. All datasets should be available in the console below:

```{r choose_your_own, exercise=TRUE}

```
