---
title: "Exploring Categorical Data"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(openintro)
library(tidyverse)

covid <- read_csv("https://raw.githubusercontent.com/EricFriedlander/MATH221SP22/main/Datasets/simpsons_paradox_covid.csv")

gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```

## COVID Vaccine and Simpson's Paradox

In this activity we're going to use contingency tables and bar charts to explore dataset on Delta Variant COVID-19 cases in the UK. Let start by taking a look at the dataset. It's called `covid`. How many rows and columns does it have? What are the variables?

```{r covid-explore, exercise=TRUE}

```

When beginning to work with a dataset it makes sense to explore some of the properties of the data before starting to make inferences. In the console below generate a contingency table with `age_group` in the columns and `outcome` in the rows. To do this:

1. Pipe `covid` into the `count` function making sure to add `age_group` and `outcome` as arguments in `count`.
2. Pipe the result into the `pivot_wider` function. Make sure to use the `id_cols` and `names_from` arguments to specifiy which variables your need and which variable will be the columns. In addition, make sure you specify `values_from = n` so R knows where to get the counts from.

```{r age-cont, exercise=TRUE}

```

```{r age-cont-hint-1}
covid %>%
  _____(____, ____) %>%
  ____(id_cols = c(____, ____, ____), names_from = _____, values_from = _____)
```

```{r age-cont-hint-2}
covid %>%
  count(_____, _____) %>%
  pivot_wider(id_cols = c(_____, _____, _____), names_from = _____, values_from = _____)
```

```{r age-cont-hint-3}
covid %>%
  count(_____, _____) %>%
  pivot_wider(id_cols = c(_____, _____, n), names_from = _____, values_from = _____)
```

```{r age-cont-hint-4}
covid %>%
  count(outcome, age_group) %>%
  pivot_wider(id_cols = c(age_group, outcome, n), names_from = age_group, values_from = n)
```


Unless you need to actual numbers, a better way to summarize your data is visually, using a bar plot. Take a minute and brain storm what type of bar plot would be best to visualize this data if we're only interested in the relative counts.

```{r best-bar}
question_radio("Which plot would be best?",
               answer("Stacked Bar", message="This one would be ok but it can sometimes be hard to see what raw counts are with a stacked bar plot."),
               answer("Dodged Bar Chart", correct=TRUE),
               answer("Standardized Bar Chart", message="This isn't good since we can't see the counts in this plot."),
               allow_retry = TRUE)
```

In the console below create a dodged bar plot with `age_group` on the x-axis and the colors corresponding to `outcome`.

1. Create a plot using the `ggplot` function. Make sure to specify the dataset, the `x` variable, and `fill` variable (`outcome`)
2. Add a `geom_bar` with `position = "dodge"`.

```{r dodge-plot, exercise=TRUE}

```

```{r dodge-plot-hint-1}
ggplot(____, ____(x =_____, fill = _____)) +
  geom_bar(____ = "____")
```

```{r dodge-plot-hint-2}
ggplot(covid, aes(x =_____, fill = _____)) +
  geom_bar(fill = "____")
```

```{r dodge-plot-hint-3}
ggplot(covid, aes(x=age_group, fill=outcome)) +
  geom_bar(position = "dodge")
```

It's difficult to see the deaths on this plot. Instead, let's use faceting to visualize the survivors and dead in different plots. In the console below use the function `facet_wrap` to create bar charts for each `outcome` with `age_group` on the x-axis. To do this:

1. Create a plot using ``ggplot``, specifying the x-variable in the `aes` function. There's no need for a `fill` argument.
2. Add a `geom_bar` layer.
3. Add a `facet_wrap` layer making sure to specify `~outcome`.

```{r facet-age, exercise=TRUE}
# Create plot
____ +
  # geom_bar layer
  ____ +
  # facet wrap layer, not that the scales argument will allow the bar charts to have different y limits
  facet_wrap(____, scales="free_y")
```

```{r facet-age-hint-1}
# Create plot
ggplot(____, aes(x=____)) +
  # geom_bar layer
  geom_bar() +
  # facet wrap layer, not that the scales argument will allow the bar charts to have different y limits
  facet_wrap(____, scales="free_y")
```

```{r facet-age-hint-2}
# Create plot
ggplot(____, aes(x=____)) +
  # geom_bar layer
  geom_bar() +
  # facet wrap layer, not that the scales argument will allow the bar charts to have different y limits
  facet_wrap(~outcome, scales="free_y")
```

```{r facet-age-hint-3}
# Create plot
ggplot(covid, aes(x=age_group)) +
  # geom_bar layer
  geom_bar() +
  # facet wrap layer, not that the scales argument will allow the bar charts to have different y limits
  facet_wrap(~outcome, scales="free_y")
```

What do you notice in these plots?

### Vaccine efficacy

We'd like to explore how effective COVID-19 vaccines are at preventing death. To do this, we'll use a contingency table. Suppose we split out the rows by `vaccine_status` and the columns by `outcome`.

```{r rowsum-q}
question_radio("What type of table will be best to assess the efficacy of the vaccine?",
               answer("Counts"),
               answer("Total proportions"),
               answer("Proportions, rows sum to 1", correct=TRUE),
               answer("Proportions, columns sum to 1"),
               allow_retry = TRUE
)
```

In the console below, create the contingency table discussed above. Make sure to `group_by` the correct variable, then use `mutate` to compute the proportions:

```{r cont-vax-eff, exercise=TRUE}

```

```{r cont-vax-eff-hint-1}
covid %>%
  group_by(____) %>%
  count(____) %>%
  mutate(prop = ____) %>%
  pivot_wider(id_cols=c(____, ____, _____), names_from = _____, values_from = _____)
```

```{r cont-vax-eff-hint-2}
covid %>%
  group_by(vaccine_status) %>%
  count(outcome) %>%
  mutate(prop = n/sum(n)) %>%
  pivot_wider(id_cols=c(vaccine_status, outcome, prop), names_from = outcome, values_from = prop)
```

Based on these results what are your initial impressions on the efficacy of the vaccine in preventing death due to the delta variant?

### Simpson's Paradox

In the console below, create a standardized bar chart showing the proportion of vaccinated and unvaccinated individuals who from each age group. You can use similar to code to above, just make sure you specify `position="fill"` in the `geom_bar` layer. I encourage you to also label the y-axis as "Proportion":

```{r stand-bar1, exercise=TRUE}

```

```{r stand-bar1-hint-1}
ggplot(____, aes(x=____, fill=____)) +
  geom_bar(____="____") +
  labs(y="Proportion")
```

```{r stand-bar1-hint-2}
ggplot(covid, aes(x=vaccine_status, fill=age_group)) +
  geom_bar(position="fill")+
  labs(y="Proportion")
```

What do you notice?

Hmmm... Let's take a look at the mortality rate broken out by `vaccine_status` and `age_group`. Create two contingency tables below, one for each age bucket. You can use the `filter(age_group == "50+")`  and `filter(age_group == "under 50")` functions to help you (note that for the younger age group it will round some numbers to 1):

```{r cavid-by-age, exercise=TRUE}
# 50+ table
____ %>%
  filter(____) %>%
  group_by(____) %>%
  count(____) %>%
  mutate(____) %>%
  pivot_wider(id_cols=c(_____, _____, _____), names_from = _____, values_from = _____)

# under 50 table (you can use very similar code to the table above)

```

```{r cavid-by-age-solution}
# 50+ table
covid %>%
  filter(age_group == "50 +") %>%
  group_by(vaccine_status) %>%
  count(outcome) %>%
  mutate(prop = n/sum(n)) %>%
  pivot_wider(id_cols=c(vaccine_status, outcome, prop), names_from = outcome, values_from = prop)

# under 50 table (you can use very similar code to the table above)
covid %>%
  filter(age_group == "under 50") %>%
  group_by(vaccine_status) %>%
  count(outcome) %>%
  mutate(prop = n/sum(n)) %>%
  pivot_wider(id_cols=c(vaccine_status, outcome, prop), names_from = outcome, values_from = prop)
```

Based on these results have your impressions on the efficacy of the vaccine in preventing death due to the delta variant changed? This reversal in trend is called **Simpson's Paradox**. How is it possible? What is driving this reversal?


## Heart transplant data display

This is the beginning of your homework. Please submit all answers into this [google form.](https://forms.gle/CwbADH5KEKtD8fnn8)

The Stanford University Heart Transplant Study was conducted to determine whether an experimental heart transplant program increased lifespan. Each patient entering the program was officially designated a heart transplant candidate, meaning that he was gravely ill and might benefit from a new heart. Patients were randomly assigned into treatment and control groups. Patients in the treatment group received a transplant, and those in the control group did not. The visualization below displays two different versions of the data. (Turnbull, Brown, and Hu 1974)

The data from this study is stored in the data frame `heart_transplant`. In the console below, create a contingency table the columns represent the outcome of the transplant (i.e. `alive` or `dead`) and the rows represent whether or not the patient received a transplant. Note that you will have figure out what the numbers of the variables are.

```{r contingency-table, exercise=TRUE}

```

```{r contingency-table-hint-1}
# You must use the count function and the pivot wider funder
```

```{r contingency-table-hint-2}
heart_transplant %>%
  count(_____, _____) %>%
  pivot_wider(names_from = _____, values_from = _____)
```

```{r contingency-table-hint-3}
heart_transplant %>%
  count(survived, transplant) %>%
  pivot_wider(names_from = transplant, values_from = n)
```

```{r contingency-table-check}
grade_result(
  pass_if(~identical(.result, heart_transplant %>%
  count(survived, transplant) %>%
  pivot_wider(names_from = transplant, values_from = n)))
)

```

Recreate the same contingency table as above, except now make the entries proportions:
```{r prop-table, exercise=TRUE}

```

```{r prop-table-hint-1}
# You can largely use the same code as above with an extra mutate to create the proportions and you must add an id_cols argument to your pivot_wider function to select the survived, transplant, and proportion columns (note the proportion columns will be called hwatever you name it in your mutate function)
```

```{r prop-table-hint-2}
heart_transplant %>%
  count(_____, _____) %>%
  mutate(_____, _____) %>%
  pivot_wider(id_cols = c(_____, _____, _____), names_from = _____, values_from = _____)
```

```{r prop-table-hint-3}
heart_transplant %>%
  count(survived, transplant) %>%
  mutate(prop = n/sum(n)) %>%
  pivot_wider(id_cols = c(survived, transplant, prop), names_from = transplant, values_from = prop)

```

```{r prop-table-check}
grade_result(
  pass_if(~identical(.result, heart_transplant %>%
  count(survived, transplant) %>%
  mutate(prop = n/sum(n)) %>%
  pivot_wider(id_cols = c(survived, transplant, prop), names_from = transplant, values_from = prop)
))
)

```

Recreate this contingency tables with the row proportions:
```{r rowprop-table, exercise=TRUE}

```

```{r rowprop-table-hint-1}
# You can largely use the same code as above adding in a group_by command to group by the survived variable
```

```{r rowprop-table-hint-2}
heart_transplant %>%
  count(_____, _____) %>%
  group_by(______) %>%
  mutate(_____, _____) %>%
  pivot_wider(id_cols = c(_____, _____, _____), names_from = _____, values_from = _____)
```

```{r rowprop-table-hint-3}
heart_transplant %>%
  count(survived, transplant) %>%
  group_by(survived) %>%
  mutate(prop = n/sum(n)) %>%
  pivot_wider(id_cols = c(survived, transplant, prop), names_from = transplant, values_from = prop)

```

```{r rowprop-table-check}
grade_result(
  pass_if(~identical(.result, heart_transplant %>%
  count(survived, transplant) %>%
  group_by(survived) %>%
  mutate(prop = n/sum(n)) %>%
  pivot_wider(id_cols = c(survived, transplant, prop), names_from = transplant, values_from = prop)
))
)
```

Recreate this contingency tables with the columns proportions:
```{r colprop-table, exercise=TRUE}

```

```{r colprop-table-hint-1}
# You can largely use the same code as above but simply change the variable you group by
```

```{r colprop-table-hint-2}
heart_transplant %>%
  count(_____, _____) %>%
  group_by(______) %>%
  mutate(_____, _____) %>%
  pivot_wider(id_cols = c(_____, _____, _____), names_from = _____, values_from = _____)
```

```{r colprop-table-hint-3}
heart_transplant %>%
  count(survived, transplant) %>%
  group_by(transplant) %>%
  mutate(prop = n/sum(n)) %>%
  pivot_wider(id_cols = c(survived, transplant, prop), names_from = transplant, values_from = prop)

```

```{r colprop-table-check}
grade_result(
  pass_if(~identical(.result, heart_transplant %>%
  count(survived, transplant) %>%
  group_by(transplant) %>%
  mutate(prop = n/sum(n)) %>%
  pivot_wider(id_cols = c(survived, transplant, prop), names_from = transplant, values_from = prop)
))
)
```



```{r survival-question}
question_radio("If you are interested on whether receiving a transplant increases a patients change of survival, which table would be most useful?",
               answer("Contingency table of counts"),
               answer("Contingency table of proportions"),
               answer("Contingency table of row proportions"),
               answer("Contingency table of columns proportions", correct=TRUE, message="You'd want columnn proportions since it shows you whether a higher proportion of patients survive after receiving a transplant."),
               allow_retry=TRUE
)
```

It's frequently difficult to read data in a table like this. Let's put it in a plot. In the console below create a standardized bar chart where each bar represents whether they received a treatment and the colors represent whether they survived. Make sure to rename the y-axis label to "Proportion":

```{r standardized-plot, exercise=TRUE}

```

```{r standardized-plot-hint-1}
ggplot(_____, _____(x=_____, fill=_____)) +
  geom_bar(_____ = "______") +
  labs(y = "_____")
```

```{r standardized-plot-hint-2}
ggplot(_____, aes(x=_____, fill=_____)) +
  geom_bar(position = "______")+
  labs(y = "_____")
```

```{r standardized-plot-hint-3}
ggplot(heart_transplant, aes(x=transplant, fill=survived)) +
  geom_bar(position = "fill") +
  labs(y = "Position")
```


Does receiving a transplant seem to impact survival? Briefly explain.

```{r association, exercise=TRUE}
#
```

```{r association-solution}
# Yes, from the plot it is clear that a smalle proportion of partients die after receiving a transplant.
```