---
title: "13 Inference With Mathematical Models"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(openintro)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


## Computing Probabilities with the Normal Distribution

Suppose that the weight of newborn babies is normally distributed with mean $\mu=7.5$ lb and standard deviation $0.7$ lb.

For each of the questions below, generate a plot shading the region of a normal curve corresponding to the requested probability and compute the true probability.

What is the probability of a baby weighing less than 7.5 lb? (You should be able to do this in your head). 

```{r prob-1, exercise=TRUE}

```

What is the probability of a baby weighing less than 6.5 lb?

```{r prob-2, exercise=TRUE}

```

What is the probability that a new born weighs at least 8 lb?

```{r prob-3, exercise=TRUE}

```

What is the probability that a new born weights between 7 and 9 lb?

```{r prob-4, exercise=TRUE}

```

What is the probability that a new born weights between 5.5 and 7 lb?

```{r prob-5, exercise=TRUE}

```

What is the probability that a new born weight more than 8.5 lb or less than 5.5 lb?

```{r prob-6, exercise=TRUE}

```

What is the probability that a new born weight more than 9 lb or less than 6 lb?

```{r prob-7, exercise=TRUE}

```

What is the probability that a new born weight more than 9 lb or less than 5 lb?

```{r prob-8, exercise=TRUE}

```

## Computing Quantiles/Percentiles with the Normal Distribution

Suppose that the weight of newborn babies is normally distributed with mean $\mu=7.5$ lb and standard deviation $0.7$ lb.

For each of the questions below, generate a plot shading the region of a normal curve corresponding to the requested probability and compute the requested quantities.

What is the 90th percentile of baby weights?

```{r perc-1, exercise=TRUE}

```

Doctors recommend bed rest for pregnant women whose babies are projected to weigh less than 95% of newborns. What is this weight? (Note: This is a made up fact)

```{r perc-2, exercise=TRUE}

```

Doctors recommend a c-section for women whose babies are projected to weigh more than 96.5% of newborns. What is this weight? (Note: This is a made up fact)

```{r perc-3, exercise=TRUE}

```

A baby is considered a "normal weight" if it fall in the middle 90% of weights. What are the cutoffs for this interval? (Note: This is a made up fact)

```{r perc-4, exercise=TRUE}

```

Doctors recommend extra screenings for babies predicted to be in the 20% of baby weights that are farthest away from the mean (7.5 lb). What are the cutoffs for the extra monitoring?

```{r perc-5, exercise=TRUE}

```
