---
title: "13 Inference With Mathematical Models"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(openintro)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


## Computing Probabilities with the Normal Distribution

Suppose that the weight of newborn babies is normally distributed with mean $\mu=7.5$ lb and standard deviation $0.7$ lb.

For each of the questions below, generate a plot shading the region of a normal curve corresponding to the requested probability and compute the true probability.

What is the probability of a baby weighing less than 7.5 lb? (You should be able to do this in your head). 

```{r prob-1, exercise=TRUE}

```

What is the probability of a baby weighing less than 6.5 lb?

```{r prob-2, exercise=TRUE}

```

What is the probability that a new born weighs at least 8 lb?

```{r prob-3, exercise=TRUE}

```

What is the probability that a new born weights between 7 and 9 lb?

```{r prob-4, exercise=TRUE}

```

What is the probability that a new born weights between 5.5 and 7 lb?

```{r prob-5, exercise=TRUE}

```

What is the probability that a new born weight more than 8.5 lb or less than 5.5 lb?

```{r prob-6, exercise=TRUE}

```

What is the probability that a new born weight more than 9 lb or less than 6 lb?

```{r prob-7, exercise=TRUE}

```

What is the probability that a new born weight more than 9 lb or less than 5 lb?

```{r prob-8, exercise=TRUE}

```

## Computing Quantiles/Percentiles with the Normal Distribution

Suppose that the weight of newborn babies is normally distributed with mean $\mu=7.5$ lb and standard deviation $0.7$ lb.

For each of the questions below, generate a plot shading the region of a normal curve corresponding to the requested probability and compute the requested quantities.

What is the 90th percentile of baby weights?

```{r perc-1, exercise=TRUE}

```

Doctors recommend bed rest for pregnant women whose babies are projected to weigh less than 95% of newborns. What is this weight? (Note: This is a made up fact)

```{r perc-2, exercise=TRUE}

```

Doctors recommend a c-section for women whose babies are projected to weigh more than 96.5% of newborns. What is this weight? (Note: This is a made up fact)

```{r perc-3, exercise=TRUE}

```

A baby is considered a "normal weight" if it fall in the middle 90% of weights. What are the cutoffs for this interval? (Note: This is a made up fact)

```{r perc-4, exercise=TRUE}

```

Doctors recommend extra screenings for babies predicted to be in the 20% of baby weights that are farthest away from the mean (7.5 lb). What are the cutoffs for the extra monitoring?

```{r perc-5, exercise=TRUE}

```

## Twitter users and news. 

A poll conducted in 2013 found that 52% of U.S. adult Twitter users get at least some news on Twitter. The standard error for this estimate was 2.4%, and a normal distribution may be used to model the sample proportion. (Pew Research Center 2013b)

a) Construct a 99% confidence interval for the fraction of U.S. adult Twitter users who get some news on Twitter, and interpret the confidence interval in context.

b) Identify each of the following statements as true or false. Provide an explanation to justify each of your answers.

    i) The data provide statistically significant evidence that more than half of U.S. adult Twitter users get some news through Twitter. Use a significance level of $\alpha=0.01

    ii) Since the standard error is 2.4%, we can conclude that 97.6% of all U.S. adult Twitter users were included in the study.

    iii) If we want to reduce the standard error of the estimate, we should collect less data.

    iv) If we construct a 90% confidence interval for the percentage of U.S. adults Twitter users who get some news through Twitter, this confidence interval will be wider than a corresponding 99% confidence interval.
    
```{r hw-1, exercise=TRUE}
# a)

# b)

# i)

# ii)

# iii)

# iv)
```
    
```{r hw-1-solution}
# a) Recall that the general formula is point estimate ± z⋆×SE. First, identify the three different values. The point estimate is 45%, z⋆=2.58 for a 99% confidence level, and SE=2.4%. Then, plug the values into the formula:
# 52%±2.58×2.4% → (45.8%,58.2%)
# We are 99% confident that 45.8% to 58.2% of U.S. adult Twitter users get some news on Twitter.

# b)

# i) False. 50% is included in the 99% confidence interval, hence a null hypothesis of p=0.50 would not be rejected at this level.

# ii) False. The standard error measures the variability of the sample proportion, and is unrelated to the proportion of the population included in the study.

# iii) False. We need to increase the sample size to decrease the standard error.

# iv) False. As the confidence level decreases so does the margin of error, and hence the width of the confidence interval.
```

## Mental health. 

The General Social Survey asked the question: "For how many days during the past 30 days was your mental health, which includes stress, depression, and problems with emotions, not good?" Based on responses from 1,151 US residents, the survey reported a 95% confidence interval of 3.40 to 4.24 days in 2010.

a) Interpret this interval in context of the data.

b) What does "95% confident" mean? Explain in the context of the application.

c) Suppose the researchers think a 99% confidence level would be more appropriate for this interval. Will this new interval be smaller or wider than the 95% confidence interval?

d) If a new survey were to be done with 500 Americans, do you think the standard error of the estimate be larger, smaller, or about the same.

```{r hw-2, exercise=TRUE}
# a)

# b)

# c)

# d)
```

```{r hw-2-solution}
# a) We are 90% confident that US residents experience poor mental health 3.40 to 4.24 days per month.

# b) 90% of random samples of size 1,151 will yield a confidence interval that contains the true average number of bad mental health days that US residents experience per month.

# c) To be more sure they capture the actual mean, they require a wider interval, unless they collect more data.
# 
# d) Less data means less precision. The estimate will probably be less accurate with less data, so the interval will be larger.
```

## Repeated student samples. 

Of all freshman at a large college, 16% made the dean’s list in the current year. As part of a class project, students randomly sample 40 students and check if those students made the list. They repeat this 1,000 times and build a distribution of sample proportions.

a) What is this distribution called?

b) Would you expect the shape of this distribution to be symmetric, right skewed, or left skewed? Explain your reasoning.

c) What is the name of the variability of this distribution?

d) Suppose the students decide to sample again, this time collecting 90 students per sample, and they again collect 1,000 samples. They build a new distribution of sample proportions. How will the variability of this new distribution compare to the variability of the distribution when each sample contained 40 observations?

```{r hw-3, exercise=TRUE}
# a)

# b)

# c)

# d)
```

```{r hw-3-solution}
# a) Sampling distribution.
# 
# b) Since the proportion is p=0.16 and n=40, the success-failure condition is not satisfied, with the expected number of successes being just 40×0.16=6.4. When we have too few expected successes, the sampling distribution of p̂  is right-skewed.
# 
# c) Standard error.
# 
# d) When there are more observations in the sample, the point estimate tends to be less variable. This means the distribution will tend to be less variable when we have more observations per sample. Beyond the required answer: the success-failure condition would be satisfied with this larger sample, so the distribution would also be symmetric in this scenario.
```

## Submit

```{css echo=FALSE}
@media print {
  .topicsContainer,
  .topicActions,
  .exerciseActions .skip {
    display: none;
  }
  .topics .tutorialTitle,
  .topics .section.level2,
  .topics .section.level3:not(.hide) {
    display: block;
  }
  .topics {
    width: 100%;
  }
  .tutorial-exercise, .tutorial-question {
    page-break-inside: avoid;
  }
  .section.level3.done h3 {
    padding-left: 0;
    background-image: none;
  }
  .topics .showSkip .exerciseActions::before {
    content: "Topic not yet completed...";
    font-style: italic;
  }
}
```

Make sure you have the tutorial open in a browser (preferably Google Chrome). Please press `print page` button below to print the tutorial. Print it to a `pdf` file and upload it into Moodle. Make sure all code boxed are visible. If you can't do that please submit your answers to [the Google Form](https://forms.gle/13PDq4eW1j5t4tpE7).

```{js print2pdf1, context="server"}
    // the following 2 chunks print the completed sections of the tutorial to PDF
    // uses "css/print2pdf.css"
    $(document).on('shiny:inputchanged', function(event) {
      if (event.name === 'print2pdf') {
        window.print();
      }
    });
```
    
```{r print2pdf2}
    # button can be placed anywhere in the tutorial
    actionButton("print2pdf", "Print page", style="opacity: .7; color: #000;")
```

